---
title: "Comparative Data Analysis of Virtual Screening
Methodologies for Predicting Urease Inhibitory Activity"
output: html_notebook
---

### **User Configuration**

This section allows the user to specify the directory where the input
data is stored. The `method_type` parameter determines the type of
virtual screening (VS) method being processed (`Docking`, `MM-GBSA`, or
`EnsembleDocking`). If experimental IC₅₀ values are stored in a separate
file, the path to this file should be provided.

Notate that Docking (standard) has a similar output to Induced-Fit
Docking and Quantum-Polarized Ligand-Docking.

```{r}

# Load required libraries
library(openxlsx)  # For reading/writing Excel files
library(dplyr)     # For data manipulation

# -------------------------
# User Configuration
# -------------------------

# Define the path to the input files and method type
path <- "path/to/your/data"  # Set your working directory
method_type <- "Docking"  # Options: "Docking", "MMGBSA", "EnsembleDocking"
ic50_file <- "IC50_data.csv"  # If available, path to IC50 data



```

### **Loading Data**

-   If `method_type = "EnsembleDocking"`, multiple CSV files are loaded
    and combined into a single dataset.

-   The script extracts frame numbers from the filenames to properly
    annotate the dataset.

-   Assumes filenames follow the format \*\_n.csv where n is the frame
    number.

-   If `method_type` is `Docking` or `MM-GBSA`, a single input file is
    read, either in CSV or XLSX format.

```{r}
# -------------------------
# Load Data
# -------------------------

if (method_type == "EnsembleDocking") {
  # Get a list of all CSV files in the directory
  # This is essential for ensemble docking, where multiple frames are analyzed.
  file_list <- list.files(path, pattern = "*.csv", full.names = TRUE)
  
  # Initialize an empty list to store data
  data_list <- list()
  
  # Loop through each CSV file to extract relevant data
  for (file_path in file_list) {
    data <- read.csv(file_path, stringsAsFactors = FALSE)
    
    # Extract Frame number from filename
    # Assumes filenames follow the format *_n.csv where n is the frame number.
    frame_number <- gsub(".*_SP_(\\d+)\\.csv", "\\1", basename(file_path))
    data$Frame <- as.numeric(frame_number)
    
    data_list[[length(data_list) + 1]] <- data
  }
  
  # Combine all extracted data into a single dataframe
  data <- bind_rows(data_list)
  
} else {
  # Single file processing for Docking/MMGBSA
  # The script expects either a CSV or XLSX file containing processed docking or MMGBSA data.
  file_name <- "input_file.csv"  # Modify based on actual data
  file_path <- file.path(path, file_name)
  
  if (grepl(".csv$", file_name, ignore.case = TRUE)) {
    data <- read.csv(file_path, stringsAsFactors = FALSE)
  } else if (grepl(".xlsx$", file_name, ignore.case = TRUE)) {
    data <- read.xlsx(file_path, sheet = 1)
  } else {
    stop("Unsupported file format. Please use CSV or XLSX.")
  }
}

```

### **Preprocessing**

-   Removes unnecessary rows such as minimized structures that may have
    been retained in docking outputs.

-   Selects the appropriate columns depending on the method type.

-   Ensures that only available columns are included in the final
    dataset.

```{r}
# -------------------------
# Preprocessing
# -------------------------

# Remove minimized structures if present
data <- data[!grepl(".* - minimized", data$Title), ]  

# Define column names based on method type
if (method_type == "Docking" || method_type == "EnsembleDocking") {
  cols_selected <- c("Title", "docking.score", "glide.ligand.efficiency", 
                     "glide.gscore", "glide.emodel", "glide.energy", "glide.einternal", 
                     "XP.PoseRank", "IC50.uM.Exp", "pIC50.Exp", "Frame")
} else if (method_type == "MMGBSA") {
  cols_selected <- c("Title", "r_psp_MMGBSA_dG_Bind", "r_psp_MMGBSA_dG_Bind_Coulomb", 
                     "r_psp_MMGBSA_dG_Bind_Hbond", "r_psp_MMGBSA_dG_Bind_Lipo", 
                     "r_psp_MMGBSA_dG_Bind_Solv_GB", "r_psp_MMGBSA_dG_Bind_vdW", 
                     "r_psp_Prime_MMGBSA_ligand_efficiency", "IC50", "pIC50")
} else {
  stop("Invalid method type. Use 'Docking', 'MMGBSA', or 'EnsembleDocking'.")
}

# Retain only available columns
data <- data[, cols_selected, drop = FALSE]


```

### **Handling Experimental IC₅₀ Data**

-   If an external file containing IC₅₀ values is available, it is
    merged with the dataset.

-   Compounds without experimental IC₅₀ values are assigned a default
    value (`800 µM` for IC₅₀ and `3.09` for pIC₅₀) to maintain data
    consistency.

-   Converts IC₅₀ values to numeric format to ensure compatibility with
    statistical analyses.

```{r}
# -------------------------
# Handling Experimental IC50 Data
# -------------------------

if (file.exists(ic50_file)) {
  ic50_data <- read.csv(ic50_file, stringsAsFactors = FALSE)
  data$IC50 <- ic50_data$IC50.uM.Exp[match(data$Title, ic50_data$Title)]
  data$pIC50 <- ic50_data$pIC50[match(data$Title, ic50_data$Title)]
}

# Replace missing IC50 values with defaults
data$IC50 <- ifelse(is.na(data$IC50), 800, data$IC50)
data$pIC50 <- ifelse(is.na(data$pIC50), 3.09, data$pIC50)

# Convert to numeric
data$IC50 <- as.numeric(data$IC50)
data$pIC50 <- as.numeric(data$pIC50)



```

### **Docking-Specific Processing**

-   The dataset is sorted by compound name (`Title`) and docking score,
    ensuring that the best-ranked poses are prioritized.

-   Each compound is assigned a `Pose` number, indicating its ranking
    within the set of poses based on docking score.

```{r}
# -------------------------
# Docking-Specific Processing
# -------------------------
if (method_type == "Docking" || method_type == "EnsembleDocking") {
  # Sort by Title and docking score
data <- data[order(data$Title, data$docking.score), ]

  # Assign Pose numbers
data$Poses <- ave(data$docking.score, data$Title, FUN = function(x) seq_along(x))
}

```

### **MMGBSA-Specific Processing**

-   The dataset is sorted by compound name (`Title`) and MM-GBSA binding
    energy (`r_psp_MMGBSA_dG_Bind`), ensuring that the most favorable
    energy values are ranked first.

-   Each compound is assigned a `Pose` number based on MM-GBSA binding
    energy.

```{r}

# -------------------------
# MMGBSA-Specific Processing
# -------------------------
if (method_type == "MMGBSA") {
  # Sort by Title and binding energy
data <- data[order(data$Title, data$r_psp_MMGBSA_dG_Bind), ]

  # Assign Pose numbers
data$Poses <- ave(data$r_psp_MMGBSA_dG_Bind, data$Title, FUN = function(x) seq_along(x))
}


```

### **Exporting Processed Data**

-   The final processed dataset is saved as an Excel file
    (`combined_data.xlsx`) for downstream statistical analysis.

-   Displays the first few rows of the processed dataset to verify
    correctness.

```{r}
# -------------------------
# Export Processed Data
# -------------------------

output_file <- file.path(path, "combined_data.xlsx")
write.xlsx(data, output_file)

# Display first few rows
head(data)
```
