---
title: "Comparative Data Analysis of Virtual Screening
Methodologies for Predicting Urease Inhibitory Activity"
output: html_notebook
---

## IC50 verification

This segment is not necessary, build a file with the compounds, IC50,
pIC50 and physicochemical characteristics

```{r}

# Load necessary libraries
library(dplyr)
library(readr)

# -------------------------
# User Configuration
# -------------------------

# Define the file path
IC50_check <- "01_EDA/For_IC50.csv"

# -------------------------
# Step 1: Read the CSV file and identify unique Titles
# -------------------------

# Read the CSV file, automatically detecting column types
csv_data <- read_csv(IC50_check, show_col_types = FALSE)

# Ensure that the "Title" column exists before proceeding
if (!"Title" %in% colnames(csv_data)) {
  stop("Error: The column 'Title' is missing from the dataset.")
}

# -------------------------
# Step 2: Rename columns
# -------------------------

# Rename columns: "IC50-uM-Exp" -> "IC50", "pIC50" -> "pIC50.Exp"
csv_data <- csv_data %>%
  rename(
    IC50 = `IC50-uM-Exp`
    #pIC50.Exp = pIC50
  )

# -------------------------
# Step 3: Select only the desired columns
# -------------------------

# Define the columns to keep
selected_columns <- c(
  "Title", "#stars", "#amine", "#amidine", "#acid", "#amide", "#rotor", "rtvFG", "CNS", "mol MW", "dipole",
  "SASA", "FOSA", "FISA", "PISA", "WPSA", "volume", "donorHB", "accptHB", "dip^2/V", "ACxDN^.5/SA", "glob",
  "QPpolrz", "QPlogPC16", "QPlogPoct", "QPlogPw", "QPlogPo/w", "QPlogS", "CIQPlogS", "QPlogHERG", "QPPCaco",
  "QPlogBB", "QPPMDCK", "QPlogKp", "IP(eV)", "EA(eV)", "#metab", "QPlogKhsa", "HumanOralAbsorption",
  "PercentHumanOralAbsorption", "SAfluorine", "SAamideO", "PSA", "#NandO", "RuleOfFive", "RuleOfThree",
  "#ringatoms", "#in34", "#in56", "#noncon", "#nonHatm", "IC50", "pIC50"
)

# Ensure only available columns are selected (avoids errors if some are missing)
available_columns <- intersect(selected_columns, colnames(csv_data))
csv_data_filtered <- csv_data %>% select(all_of(available_columns))

# -------------------------
# Step 4: Keep unique Titles and store the final dataframe
# -------------------------

# Keep only unique rows based on Title while retaining all selected columns
IC50_data <- csv_data_filtered %>% distinct(Title, .keep_all = TRUE)

# Remove rows where Title ends with " - minimized"
IC50_data <- IC50_data %>% filter(!grepl(" - minimized$", Title))

# Ensure IC50 is in the second position after Title
IC50_data <- IC50_data %>% select(Title, IC50, pIC50, everything())

# Calculate pIC50 and place it as the third column
#IC50_data <- IC50_data %>%
#  mutate(pIC50 = -log10(IC50 * 1e-6)) %>%
#  select(Title, IC50, pIC50, everything())  # Ensure correct column order

# Save the processed dataset as a CSV file
#write_csv(IC50_data, "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/EDA/IC50_data_cleaned_EDA.csv")





```

### Dont use!, just count the unique compounds

```{r}
# Load necessary libraries
library(dplyr)
library(readr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the CSV files
data_dir <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/EDA/1/"

# Get a list of all CSV files in the directory
csv_files <- list.files(path = data_dir, pattern = "*.csv", full.names = TRUE)

# -------------------------
# Read CSV files and count unique Title values with valid IC50
# -------------------------

# Function to count unique Titles with valid IC50 values
count_filtered_unique_titles <- function(file) {
  data <- read_csv(file, show_col_types = FALSE)  # Read CSV file
  
  # Check if required columns exist
  if (!("Title" %in% colnames(data)) || !("IC50-uM-Exp" %in% colnames(data))) {
    warning(paste("Warning: Required columns not found in", file))
    return(data.frame(File = basename(file), Unique_Titles = NA))
  }
  
  # Convert IC50 column to numeric (in case of character format)
  data <- data %>%
    mutate(`IC50-uM-Exp` = as.numeric(`IC50-uM-Exp`)) %>%
    filter(!is.na(`IC50-uM-Exp`))  # Remove NA values in IC50
  
  # Count unique values in Title after filtering
  unique_count <- length(unique(data$Title))
  
  return(data.frame(File = basename(file), Unique_Titles = unique_count))
}

# Apply function to each CSV file and combine results
filtered_unique_counts <- bind_rows(lapply(csv_files, count_filtered_unique_titles))

# Display the summary table
print(filtered_unique_counts)




```

## EDA

### Dataset

We use different sets of coumarins molecules already reported (CIAR) as
inhibitors for urease. These molecules were collected manually from:

-   Paper1: Bis-coumarins; Non-Cytotoxic Selective Urease Inhibitors and
    Antiglycation Agents. DOI:
    <https://doi.org/10.1016/j.bioorg.2019.103170>

-   Paper4: Synthesis of 2-Acylated and Sulfonated 4-hydroxycoumarins:
    In vitro Urease Inhibition and Molecular Docking Studies. DOI:
    <http://dx.doi.org/10.1016/j.bioorg.2016.04.005>

-   Paper6: Design, synthesis, in vitro evaluation, molecular docking
    and ADME properties studies of hybrid bis-coumarin with thiadiazole
    as a new inhibitor of Urease. DOI:
    <https://doi.org/10.1016/j.bioorg.2019.103235>

-   Paper7: 4-Oxycoumarinyl Linked Acetohydrazide Schiff bases as Potent
    Urease Inhibitors. DOI:
    <https://doi.org/10.1016/j.bioorg.2020.104365>

-   Paper8: Exploring biological efficacy of coumarin clubbed
    thiazolo[3,2--b][1,2,4]tria- zoles as efficient inhibitors of
    urease: A biochemical and in silico approach. DOI:
    <https://doi.org/10.1016/j.ijbiomac.2019.09.105>

-   Paper10: Biscoumarin: new class of urease inhibitors; economical
    synthesis and activity. DOI: <doi:10.1016/j.bmc.2004.01.010>

### Integrate inactive compounds

```{r}

IC50_data$IC50 <- ifelse(is.na(IC50_data$IC50), 800, IC50_data$IC50)
IC50_data$pIC50 <- ifelse(is.na(IC50_data$pIC50), 3.09, IC50_data$pIC50)

IC50_data$IC50 <- as.numeric(IC50_data$IC50)
IC50_data$pIC50 <- as.numeric(IC50_data$pIC50)
```

### Divide inhibitors (active) and non-inhibitors (inactive) by 50 uM

```{r}
library(dplyr)

#IC50_data <- IC50_data %>%
#  mutate(activity =
#                     case_when(IC50 < 50 ~ "UI", 
#                               IC50 >= 50 ~ "nUI",
#                               ))

```

```{r}
library(dplyr)

IC50_data <- IC50_data %>%
  mutate(activity = case_when(
    IC50 < 50 ~ "UI", 
    IC50 >= 50 & IC50 < 800 ~ "nUI",
    IC50 >= 800 ~ "ND"
  ))

```

### Histogram IC50

#### UI and nUI (added to the ND)

```{r}

library(ggplot2)

IC <- ggplot(IC50_data, aes(x = IC50)) +
geom_histogram(fill="#69b3a2", color="black", alpha=0.9) +
#geom_histogram(data=coumarin_dataset, aes(x = coumarin_dataset$pIC50, fill = coumarin_dataset$pIC50)) +
theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    axis.ticks = element_line(colour = "black", size = 0.5),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  ) + 
labs(x = "IC50 (uM)", y = "Count")

#IC
```

#### UI, nUI and ND

```{r}
library(ggplot2)

pIC <- ggplot(IC50_data, aes(x = pIC50)) +
geom_histogram(fill="#69b3a2", color="black", alpha=0.9) +
#geom_histogram(data=coumarin_dataset, aes(x = coumarin_dataset$pIC50, fill = coumarin_dataset$pIC50)) +
theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    axis.ticks = element_line(colour = "black", size = 0.5),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  ) + 
labs(x = "pIC50", y = "Count")

#pIC

```

```{r}
library(gridExtra)
Hist <- grid.arrange(IC, pIC, nrow=2)

Hist

#ggsave("Hist_IC50.tiff", plot = Hist,  width = 6.7, height = 5, dpi = 600, device = "tiff")
```

### Correlation heatmap

| Rule | Key Parameters |
|------|----------------|

|                             |                                             |
|----------------------------|-------------------------------------------|
| **Lipinski's Rule of Five** | `mol MW`, `QPlogPo/w`, `donorHB`, `accptHB` |

|                  |                                       |
|------------------|---------------------------------------|
| **Veber's Rule** | `#rotor`, `PSA`, `donorHB`, `accptHB` |

|                      |                                             |
|----------------------|---------------------------------------------|
| **Jorgensen's Rule** | `QPlogPo/w`, `QPlogS`, `QPPCaco`, `QPlogBB` |

|                      |        |
|----------------------|--------|
| **Activity Metrics** | `IC50` |

```{r}

# Load necessary libraries
library(ggcorrplot)
library(ggplot2)
library(dplyr)

# -------------------------
# Select Key Features for Correlation Analysis
# -------------------------

# Define the selected properties for Lipinski, Veber, and Jorgensen rules
selected_features <- c( "IC50",  "mol MW", "QPlogPo/w", "donorHB", "accptHB", 
                       "#rotor", "PSA", "QPlogS", "QPPCaco", "QPlogBB")

# Ensure only available columns are selected (avoid errors if some are missing)
available_features <- intersect(selected_features, colnames(IC50_data))

# Filter dataset to include only selected properties
filtered_data <- IC50_data %>% select(all_of(available_features))

# Compute correlation matrix
cor_matrix <- cor(filtered_data, use = "complete.obs")

# -------------------------
# Generate Improved Correlation Heatmap
# -------------------------

# Create a cleaner heatmap with better legend formatting
correlation_heatmap <- ggcorrplot(cor_matrix, method = "square", type = "lower",
           lab = TRUE, lab_size = 3,  # Adjust correlation number size
           show.legend = TRUE,
           colors = c("blue", "white", "red"),
           title = "") +
  theme(
    legend.position = "right",  # Move legend to the right
    legend.direction = "vertical",  # Make legend vertical
    legend.key.width = unit(0.2, "cm"),  # Narrower legend bar
    legend.key.height = unit(1, "cm"),  # Taller legend bar
    legend.title = element_text(size = 12, face = "bold"),  # Adjust legend title size
    legend.text = element_text(size = 12),  # Adjust legend label size
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  ) +
  guides(fill = guide_colorbar(title = "", title.position = "top"))



correlation_heatmap

#ggsave("correlation_heatmap.tiff", plot = correlation_heatmap,  width = 6.7, height = 5, dpi = 600, device = "tiff")




```

### pIC50 vs ROF

#### UI and nUI (added to the ND)

```{r}
library(ggplot2)
library(dplyr)

# Ensure 'activity' is a factor
IC50_data$activity <- factor(IC50_data$activity, levels = c("UI", "nUI"))

# Generate improved boxplot
enzimatic_activity <- ggplot(IC50_data, aes(x = factor(RuleOfFive), y = pIC50, fill = activity)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Reduce outlier visibility
  #geom_jitter(aes(color = activity), width = 0.2, alpha = 0.6, size = 2) +  # Add jitter points
  scale_fill_manual(values = c("UI" = "red", "nUI" = "blue")) +  # Custom colors
  #scale_color_manual(values = c("UI" = "darkred", "nUI" = "darkblue")) +  # Jitter color match
  labs(title = "",
       x = "Rule of Five Violations",
       y = "pIC50",
       fill = "Activity") +  # Legend title
  theme_minimal() +
  theme(
    legend.position = "right",  # Move legend above the plot
    legend.direction = "vertical",  # Make legend vertical
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

enzimatic_activity

#ggsave("Activity2.tiff", plot = enzimatic_activity,  width = 6.7, height = 5, dpi = 600, device = "tiff")

```

#### UI, nUI and ND

```{r}
library(ggplot2)
library(dplyr)

# Ensure activity is categorized correctly, including ND
IC50_data <- IC50_data %>%
  mutate(activity = factor(case_when(
    IC50 < 50 ~ "UI",
    IC50 >= 50 & IC50 < 800 ~ "nUI",
    IC50 >= 800 ~ "ND"
  ), levels = c("UI", "nUI", "ND")))  # Ensure ND appears in legend

# Generate improved boxplot including ND
enzimatic_activity <- ggplot(IC50_data, aes(x = factor(RuleOfFive), y = pIC50, fill = activity)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Reduce outlier visibility
  #geom_jitter(aes(color = activity), width = 0.2, alpha = 0.6, size = 2) +  # Add jitter points
  scale_fill_manual(values = c("UI" = "red", "nUI" = "blue", "ND" = "green")) +  # Add ND color
  #scale_color_manual(values = c("UI" = "darkred", "nUI" = "darkblue", "ND" = "black")) +  # Jitter color match
  labs(title = "",
       x = "Rule of Five Violations",
       y = "pIC50",
       fill = "Activity") +  # Legend title
  theme_minimal() +
  theme(
    legend.position = "right",  # Move legend to the right
    legend.direction = "vertical",  # Make legend vertical
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

enzimatic_activity

# Save plot with high resolution
ggsave("Activity.tiff", plot = enzimatic_activity, width = 6.7, height = 5, dpi = 600, device = "tiff")

```

### PCA

#### UI and nUI (added to the ND)

```{r}
library(ggplot2)
library(ggfortify)
library(ggrepel)
library(dplyr)

# Ensure 'activity' is a factor with UI and nUI as levels
IC50_data <- IC50_data %>%
  mutate(activity = factor(activity, levels = c("UI", "nUI")))

# Select numeric columns, ensuring no zero-variance ones are included
numeric_data <- IC50_data %>% select(where(is.numeric))

# Remove zero-variance columns
zero_variance_cols <- numeric_data %>% summarise(across(everything(), var, na.rm = TRUE)) %>%
  select(where(~ . == 0)) %>% colnames()
numeric_data_filtered <- numeric_data %>% select(-all_of(zero_variance_cols))

# Perform PCA
pca_results <- prcomp(numeric_data_filtered, center = TRUE, scale. = TRUE)

# Extract loadings (variable contributions)
loadings <- as.data.frame(pca_results$rotation[, 1:2])
loadings$Variable <- rownames(loadings)

# Filter only the most important loadings (based on absolute contribution)
top_loadings <- loadings %>%
  arrange(desc(abs(PC1) + abs(PC2))) %>%
  slice(1:10)  # Show only the 15 most important variables

# Create PCA biplot
pca <- autoplot(pca_results, data = IC50_data, colour = 'activity', shape = TRUE, alpha = 0.6) +
  stat_ellipse(aes(color = activity), level = 0.95) +  # Confidence ellipses
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +  # Add arrows for loadings
  geom_text_repel(data = top_loadings, aes(x = PC1, y = PC2, label = Variable), 
                  size = 5, color = "black", max.overlaps = 10) +  # Repel overlapping labels
  scale_color_manual(values = c("UI" = "red", "nUI" = "blue")) +  # Custom colors
  scale_shape_manual(values = c("UI" = 17, "nUI" = 16)) +  # Different shapes
  labs(title = "",
       x = paste0("PC1 (", round(pca_results$sdev[1]^2 / sum(pca_results$sdev^2) * 100, 1), "%)"),
       y = paste0("PC2 (", round(pca_results$sdev[2]^2 / sum(pca_results$sdev^2) * 100, 1), "%)"),
       color = "Activity",
       shape = "Activity") +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),  
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

pca

ggsave("pca2.tiff", plot = pca,  width = 6.7, height = 5, dpi = 600, device = "tiff")
```

#### UI, nUI and ND

```{r}
library(ggplot2)
library(ggfortify)
library(ggrepel)
library(dplyr)

# Ensure 'activity' is a factor with UI, nUI, and ND as levels
IC50_data <- IC50_data %>%
  mutate(activity = factor(activity, levels = c("UI", "nUI", "ND")))

# Select numeric columns, ensuring no zero-variance ones are included
numeric_data <- IC50_data %>% select(where(is.numeric))

# Remove zero-variance columns
zero_variance_cols <- numeric_data %>% summarise(across(everything(), var, na.rm = TRUE)) %>%
  select(where(~ . == 0)) %>% colnames()
numeric_data_filtered <- numeric_data %>% select(-all_of(zero_variance_cols))

# Perform PCA
pca_results <- prcomp(numeric_data_filtered, center = TRUE, scale. = TRUE)

# Extract loadings (variable contributions)
loadings <- as.data.frame(pca_results$rotation[, 1:2])
loadings$Variable <- rownames(loadings)

# Filter only the most important loadings (based on absolute contribution)
top_loadings <- loadings %>%
  arrange(desc(abs(PC1) + abs(PC2))) %>%
  slice(1:10)  # Show only the 10 most important variables

# Create PCA biplot
pca <- autoplot(pca_results, data = IC50_data, colour = 'activity', shape = TRUE, alpha = 0.6) +
  stat_ellipse(aes(color = activity), level = 0.95) +  # Confidence ellipses
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm")), color = "black") +  # Add arrows for loadings
  geom_text_repel(data = top_loadings, aes(x = PC1, y = PC2, label = Variable), 
                  size = 5, color = "black", max.overlaps = 10) +  # Repel overlapping labels
  scale_color_manual(values = c("UI" = "red", "nUI" = "blue", "ND" = "green")) +  # Add ND (green)
  scale_shape_manual(values = c("UI" = 17, "nUI" = 16, "ND" = 15)) +  # Different shapes for ND
  labs(title = "",
       x = paste0("PC1 (", round(pca_results$sdev[1]^2 / sum(pca_results$sdev^2) * 100, 1), "%)"),
       y = paste0("PC2 (", round(pca_results$sdev[2]^2 / sum(pca_results$sdev^2) * 100, 1), "%)"),
       color = "Activity",
       shape = "Activity") +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),  
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

pca

# Save the plot as high-resolution TIFF
#ggsave("pca.tiff", plot = pca, width = 6.7, height = 5, dpi = 600, device = "tiff")

```

### ROF distribution

#### UI and nUI (added to the ND)

```{r}
library(ggplot2)
library(dplyr)

# Ensure 'activity' is a factor
IC50_data$activity <- factor(IC50_data$activity, levels = c("UI", "nUI"))

# Create bar plot with UI and nUI separation
ROF_distribution <- ggplot(IC50_data, aes(x = factor(RuleOfFive), fill = activity)) +
  geom_bar(position = "dodge", alpha = 0.8) +  # Side-by-side bars
  scale_fill_manual(values = c("UI" = "red", "nUI" = "blue")) +  # Custom colors
  labs(title = "",
       x = "Violations",
       y = "Count",
       fill = "Activity") +  # Legend title
  theme_minimal() +
  theme(
    legend.position = "right",  # Move legend to the right
    legend.direction = "vertical",  # Make legend vertical
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

ROF_distribution

ggsave("ROF_distribution2.tiff", plot = ROF_distribution,  width = 6.7, height = 5, dpi = 600, device = "tiff")
```

#### UI, nUI and ND

```{r}
library(ggplot2)
library(dplyr)

# Ensure 'activity' is a factor with UI, nUI, and ND as levels
IC50_data$activity <- factor(IC50_data$activity, levels = c("UI", "nUI", "ND"))

# Create bar plot with UI, nUI, and ND separation
ROF_distribution <- ggplot(IC50_data, aes(x = factor(RuleOfFive), fill = activity)) +
  geom_bar(position = "dodge", alpha = 0.8) +  # Side-by-side bars
  scale_fill_manual(values = c("UI" = "red", "nUI" = "blue", "ND" = "green")) +  # Add ND color
  labs(title = "",
       x = "Violations",
       y = "Count",
       fill = "Activity") +  # Legend title
  theme_minimal() +
  theme(
    legend.position = "right",  # Move legend to the right
    legend.direction = "vertical",  # Make legend vertical
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(color = "black", face = "bold", size = 12),
    axis.text.x = element_text(colour = "black", size = 12),
    axis.text.y = element_text(colour = "black", size = 12),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white")
  )

ROF_distribution

# Save the plot as high-resolution TIFF
#ggsave("ROF_distribution2.tiff", plot = ROF_distribution, width = 6.7, height = 5, dpi = 600, device = "tiff")

```

# Volume in protein frames

```{r}

library(readxl)
library(dplyr)
library(stringr)

# Excel name
file_path <- "01_EDA/Volume_Proteins.xlsx"

# Sheet names
sheets <- c("1E9Z", "1E9Y", "6QSU", "6ZJA")

# 
read_and_process_sheet <- function(sheet) {
  df <- read_excel(file_path, sheet = sheet)  
  
  # Rename columns
  colnames(df)[1] <- "Protein"
  colnames(df)[3] <- "Area"
  colnames(df)[4] <- "Volume"
  
  # Check number of columns
  if (ncol(df) < 4) {
    stop(paste("Error: La hoja", sheet, "less of 4 columns."))
  }
  
  df <- df %>%
    mutate(Protein = gsub("_chainB", "", Protein)) %>% 
    mutate(Sheet = sheet) %>%  
    mutate(
      Area = as.numeric(str_replace(Area, "(\\d{3})$", ".\\1")),   
      Volume = as.numeric(str_replace(Volume, "(\\d{3})$", ".\\1")) 
    )
  
  return(df)
}

# Read all the sheets and combine them
combined_df <- bind_rows(lapply(sheets, function(sheet) {
  tryCatch(
    read_and_process_sheet(sheet),
    error = function(e) {
      message(paste("Error al leer la hoja:", sheet, "\n", e))
      return(NULL)
    }
  )
}))

# 
head(combined_df)

# 
if (!"Frame" %in% colnames(combined_df)) {
  combined_df <- combined_df %>%
    mutate(Frame = as.character(NA))  
} else {
  combined_df <- combined_df %>%
    mutate(Frame = as.character(Frame))  
}


new_rows <- data.frame(
  Protein = c("6ZJA", "6QSU", "1E9Y", "1E9Z"),
  Frame = c("Crystal", "Crystal", "Crystal", "Crystal"),
  Area = c(379.379, 242.700, 276.773, 108.058),  
  Volume = c(300.364, 132.359, 171.188, 30.792),
  Sheet = NA_character_  # Mantener el mismo tipo de dato
)


combined_df <- bind_rows(combined_df, new_rows)


```

```{r}

library(ggplot2)
library(dplyr)
library(gridExtra)
library(gtools)

# Modify the theme to ensure the Y axis is bold
custom_theme <- theme(
  legend.position = "top",
  legend.title = element_blank(),
  legend.key.size = unit(1, "cm"),
  legend.text = element_text(color = "black", size = 18),
  axis.title.x = element_text(color = "black", face = "bold", size = 22),  # Eje X en negrita
  axis.title.y = element_text(color = "black", face = "bold", size = 22),  # Eje Y en negrita
  axis.text.x = element_text(colour = "black", face = "bold", size = 20, angle = 45, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 20),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  panel.background = element_rect(fill = "gray90", color = NA),  # Fondo gris
  panel.grid.major = element_line(color = "white", size = 0.5),  # Líneas principales en blanco
  panel.grid.minor = element_line(color = "white", size = 0.3),
  plot.title = element_text(size = 24, face = "bold")
)



  #strip.text = element_text(size = 18, face = "bold")




# Modify the line in `labs()` for the Y axis
create_plot <- function(protein_data) {
  protein_data <- protein_data %>%
    mutate(Frame = factor(Frame, 
                          levels = c("Crystal", mixedsort(setdiff(unique(Frame), "Crystal"))), 
                          ordered = TRUE))  

  ggplot(protein_data, aes(x = Frame, y = Volume, fill = Frame == "Crystal")) +
    geom_bar(stat = "identity", position = "dodge") +  
    scale_fill_manual(values = c("#009081", "black"), labels = c("Frames", "Crystal")) +  
   labs(title = unique(protein_data$Protein), x = "Frame", y = bquote(bold(Volume ~ (Å^3))), fill = "") +
    ylim(0, 800) +  
    custom_theme  # Aplicar el theme corregido
}

# Create graphs for each protein
plots <- lapply(split(combined_df, combined_df$Protein), create_plot)

# Combine all charts
Hist <- grid.arrange(grobs = plots, ncol = 1)  # Ajusta ncol según sea necesario

# Display chart combination
print(Hist)

# Save to PDF and TIFF with new Y-axis limits
#ggsave("Volume_proteins.pdf", plot = Hist, width = 16, height = 14, dpi = 300)
ggsave("Volume_proteins.tiff", plot = Hist, width = 16, height = 14, dpi = 300)
```
