---
title: "Comparative Data Analysis of Virtual Screening
Methodologies for Predicting Urease Inhibitory Activity"
output: html_notebook
---

## **Data Fusion Methods**

```{r}
# Load necessary library
library(dplyr)

# Function to calculate Median
median_value <- function(data, energy_col) {
  data %>%
    group_by(Title) %>%
    summarise(!!energy_col := median(.data[[energy_col]], na.rm = TRUE), .groups = "drop")
}

# Function to calculate Minimum Value (now considering pose)
minimum_value <- function(data, energy_col, pose) {
  data %>%
    filter(Poses == pose) %>%  
    select(Title, !!energy_col)  
}

# Function to calculate Arithmetic Mean
arithmetic_mean <- function(data, energy_col) {
  data %>%
    group_by(Title) %>%
    summarise(!!energy_col := mean(.data[[energy_col]], na.rm = TRUE), .groups = "drop")
}

# Function to calculate Geometric Mean
geometric_mean <- function(data, energy_col) {
 data %>%
   group_by(Title) %>%
   summarise(!!energy_col := sign(mean(.data[[energy_col]], na.rm = TRUE)) *
               exp(mean(log(abs(.data[[energy_col]])), na.rm = TRUE)), .groups = "drop")
}

# Function to calculate Harmonic Mean
harmonic_mean <- function(data, energy_col) {
  data %>%
    group_by(Title) %>%
    summarise(!!energy_col := sign(mean(.data[[energy_col]], na.rm = TRUE)) * 
                (n() / sum(1 / abs(.data[[energy_col]]), na.rm = TRUE)), .groups = "drop")
}

# Function to calculate Euclidean Mean
euclidean_mean <- function(data, energy_col) {
  data %>%
    group_by(Title) %>%
    summarise(
      !!energy_col := sign(mean(.data[[energy_col]], na.rm = TRUE)) * 
                      sqrt(sum((.data[[energy_col]])^2, na.rm = TRUE) / n()), 
      .groups = "drop"
    )
}
```

# **Variant 1 - 1 (Docking)**

Need modify:

-   Modify `data_path` from **Load data**:

-   Modify `energy_columns` [docking or MM-GBSA terms] from **Correlation Data Analysis**

-   Modify `output_filename` from **Correlation Data Analysis**

-   Modify `Variant` and `Method` from **Correlation Data Analysis**

### **Load data**

Load data automatically from the data_path

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "02_Data_Variants/Data_Variant_1_1_Docking/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+") 

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)
    
    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)

```

### Correlation Data Analysis

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric 
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# Define the Data Fusion Methods
# -------------------------

apply_data_fusion <- function(data, method, energy_col, pose = NULL) {
  if (method == "Median") {
    return(median_value(data, energy_col))
  } else if (method == "Minimum") {
    return(minimum_value(data, energy_col, pose))  
  } else if (method == "Arithmetic") {
    return(arithmetic_mean(data, energy_col))
  } else if (method == "Geometric") {
    return(geometric_mean(data, energy_col))
  } else if (method == "Harmonic") {
    return(harmonic_mean(data, energy_col))
  } else if (method == "Euclidean") {
    return(euclidean_mean(data, energy_col))
  } else {
    stop("Invalid Data Fusion Method")
  }
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_1_1_Docking.xlsx"

# Define the variant and method
Variant <- "1_1"
Method <- "Docking"

# Define the energy-related columns to be analyzed
energy_columns <- c("docking.score", "glide.ligand.efficiency", 
                    "glide.gscore", "glide.emodel", "glide.energy", 
                    "glide.einternal")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# -------------------------
# Create a dataframe to store the results
# -------------------------

results_df <- data.frame(Protein = character(),
                         ExpVariable = character(),
                         DataFusion = character(),
                         EnergyTerm = character(),
                         Pose = integer(),
                         Pearson = numeric(),
                         Spearman = numeric(),
                         MAE = numeric(),
                         RMSE = numeric(),
                         IRM = numeric(),  # Nueva columna para IRM
                         N = integer(),
                         Variant = character(),
                         Method = character(),
                         stringsAsFactors = FALSE)

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]

  for (fusion_method in fusion_methods) {
    
    start_pose <- ifelse(fusion_method == "Minimum", 1, 2)

    for (i in start_pose:10) {
      
      # Filter data by poses
      filtered_data <- data_protein %>%
        filter(Poses <= i)  

      for (energy_col in energy_columns) {
        
        # Apply the Data Fusion technique
        if (fusion_method == "Minimum") {
          fused_data <- apply_data_fusion(filtered_data, fusion_method, energy_col, pose = i)  
        } else {
          fused_data <- apply_data_fusion(filtered_data, fusion_method, energy_col)
        }
        
        #Detect the correct name of the merged column
        fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

        # Merge with experimental data
        merged_data <- left_join(fused_data, data_protein[, c("Title", exp_variables)], by = "Title")
        
        # Iterate over IC50 and pIC50
        for (exp_variable in exp_variables) {
          
          # Filter non-null values for calculations
          valid_data <- merged_data %>%
            filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

          n_used <- nrow(valid_data)
          
          if (n_used < 3) {
            pose_df <- data.frame(
              Protein = protein_name,  
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = NA,
              Spearman = NA,
              MAE = NA,
              RMSE = NA,
              IRM = NA,  
              N = n_used,
              Variant = Variant,  
              Method = Method,  
              stringsAsFactors = FALSE
            )
            
          } else {
            # Calculating correlations and metrics
            cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
            cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
            mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
            rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
            irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  # Nuevo cálculo de IRM
            
            # Save results to the data frame
            pose_df <- data.frame(
              Protein = protein_name,  
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = round(cor_pearson, 3),
              Spearman = round(cor_spearman, 3),
              MAE = round(mae_value, 3),
              RMSE = round(rmse_value, 3),
              IRM = round(irm_value, 3),  # Agregamos IRM
              N = n_used,
              Variant = Variant,  
              Method = Method,  
              stringsAsFactors = FALSE
            )
          }
          
          # Add results to the final dataframe
          results_df <- bind_rows(results_df, pose_df)
          
        }  # End iteration on IC50 and pIC50
      }  # End iteration on energy terms
    }  # End iteration over poses
  }  # End iteration on fusion methods
}  # End iteration on proteins

# Save results in Excel
write.xlsx(results_df, output_filename)

#Success message
cat("Data successfully saved to", output_filename, "\n")

```

# **Variant 1 - 2 (QPLD):**

### **Load data**

-   Modify `data_path` from **Load data**:

```{r}

# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/Data_Variant_1_2_QPLD/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+")

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)
    
    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Remove last "_X" from Title
    df <- df %>% mutate(Title = str_replace(Title, "_\\d+$", ""))
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)
```

### Correlation Data Analysis (up-to 100 poses)

-   Modify `energy_columns` [docking or MM-GBSA terms] from **Correlation Data Analysis**

-   Modify `output_filename` from **Correlation Data Analysis**

-   Modify `Variant` and `Method` from **Correlation Data Analysis**

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export
library(stringr)   # For regex operations

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_1_2_QPLD_100poses.xlsx"

# Define the variant and method
Variant <- "1_2"
Method <- "QPLD"

# Define the energy-related columns to be analyzed
energy_columns <- c("docking.score", "glide.ligand.efficiency", 
                    "glide.gscore", "glide.emodel", "glide.energy", 
                    "glide.einternal")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Create a dataframe to store the results
results_df <- data.frame(Protein = character(),
                         ExpVariable = character(),
                         DataFusion = character(),
                         EnergyTerm = character(),
                         Pose = integer(),
                         Pearson = numeric(),
                         Spearman = numeric(),
                         MAE = numeric(),
                         RMSE = numeric(),
                         IRM = numeric(),  # Nueva columna para IRM
                         N = integer(),
                         Variant = character(),
                         Method = character(),
                         stringsAsFactors = FALSE)

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  cat("\n🔎 Processing Protein:", protein_name, "\n")

  for (fusion_method in fusion_methods) {
    
    start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
    #max_poses <- ifelse(fusion_method == "Minimum", 10, 100)
    max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"

    for (i in start_pose:max_poses) {
      
      # Filter data by poses
      filtered_data <- data_protein %>%
        filter(Poses <= i)  

      for (energy_col in energy_columns) {
        
        # Apply the Data Fusion technique
        if (fusion_method == "Median") {
          fused_data <- median_value(filtered_data, energy_col)
        } else if (fusion_method == "Minimum") {
          fused_data <- minimum_value(filtered_data, energy_col, pose = i)
        } else if (fusion_method == "Arithmetic") {
          fused_data <- arithmetic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Geometric") {
          fused_data <- geometric_mean(filtered_data, energy_col)
        } else if (fusion_method == "Harmonic") {
          fused_data <- harmonic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Euclidean") {
          fused_data <- euclidean_mean(filtered_data, energy_col)
        } else {
          stop("Invalid Data Fusion Method")
        }
        
        # Detect the correct name of the merged column
        fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]
        
        if (length(fusion_col_name) == 0) {
          cat("Skipping", energy_col, "for", protein_name, "- Pose", i, "(not found in fused_data)\n")
          next  # Skip this iteration
        }
        
        # Merge with experimental data
        merged_data <- left_join(
          fused_data %>% distinct(Title, .keep_all = TRUE), 
          data_protein %>% distinct(Title, .keep_all = TRUE), 
          by = "Title"
        )
        
        # Identify and remove duplicate columns (with .x and .y suffixes)
        cols_to_remove <- grep("\\.y$", colnames(merged_data), value = TRUE)
        merged_data <- merged_data %>%
          select(-all_of(cols_to_remove))

        # Rename columns by removing `.x` suffixes
        colnames(merged_data) <- gsub("\\.x$", "", colnames(merged_data))

        # Verify that the merged column still exists
        if (!(energy_col %in% colnames(merged_data))) {
          cat("Skipping", energy_col, "for", protein_name, "- Pose", i, "(not found in merged_data after cleanup)\n")
          next
        }

        # Iterate over IC50 and pIC50
        for (exp_variable in exp_variables) {
          
          # Filter non-null values ​​for calculations
          valid_data <- merged_data %>%
            filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[energy_col]]))

          n_used <- nrow(valid_data)
          
          if (n_used < 3) {
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = NA,
              Spearman = NA,
              MAE = NA,
              RMSE = NA,
              IRM = NA,  # We save NA in IRM if there is not enough data
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
          } else {
            # Calculating correlations and metrics
            cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[energy_col]], use = "complete.obs")
            cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[energy_col]], use = "complete.obs", method = "spearman")
            mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[energy_col]]), na.rm = TRUE)
            rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[energy_col]])^2, na.rm = TRUE))
            irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[energy_col]])  # Nuevo cálculo de IRM
            
            # Save results to the data frame
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = round(cor_pearson, 3),
              Spearman = round(cor_spearman, 3),
              MAE = round(mae_value, 3),
              RMSE = round(rmse_value, 3),
              IRM = round(irm_value, 3),  # Agregamos IRM
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
          }
          
          # Add results to the final dataframe
          results_df <- bind_rows(results_df, pose_df)
          
        }  # End iteration on IC50 and pIC50
      }  # End iteration on energy terms

      # Free memory every 10 poses
      if (i %% 10 == 0) {
        gc()
      }
      
    }  # End iteration over poses
  }  #End iteration on fusion methods
}  # End iteration on proteins

# Save results in Excel
write.xlsx(results_df, output_filename)

#Success message
cat("Data successfully saved to", output_filename, "\n")
gc()

```

# **Variant 1 - 3 (MM-GBSA):**

### **Load data**

-   Modify `data_path` from **Load data**:

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/Data_Variant_1_3_MMGBSA/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+")

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)

    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists (it will be recalculated)
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)
```

### Correlation Data Analysis (up-to 100)

-   Modify `energy_columns` [docking or MM-GBSA terms] from **Correlation Data Analysis**

-   Modify `output_filename` from **Correlation Data Analysis**

-   Modify `Variant` and `Method` from **Correlation Data Analysis**

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proportion of predictions within the error threshold
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_1_3_MMGBSA_100poses.xlsx"

# Define the variant and method
Variant <- "1_3"
Method <- "MMGBSA"

# Define the energy-related columns to be analyzed
energy_columns <- c("MMGBSABind", "MMGBSABind_Coulomb", "MMGBSABind_Hbond", 
                    "MMGBSABind_Lipo", "MMGBSABind_Solv_GB", "MMGBSABind_vdW",
                    "Prime_MMGBSA_ligand_efficiency")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Create an empty list to store results before combining
# -------------------------

results_list <- list()  

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  
  cat("\n🔎 Processing Protein:", protein_name, "\n")
  
  # Iterate over each Data Fusion Method
  for (fusion_method in fusion_methods) {
    
    start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
    #max_poses <- ifelse(fusion_method == "Minimum", 10, 100)  # Define max poses
    max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"
    
    # Iterate over the number of poses
    for (i in start_pose:max_poses) {
      
      # Filter the dataset to include only the top `i` poses
      filtered_data <- data_protein %>%
        filter(Poses <= i)

      # Iterate over each energy-related variable
      for (energy_col in energy_columns) {
        
        # Apply the selected Data Fusion Method
        if (fusion_method == "Minimum") {
          fused_data <- minimum_value(filtered_data, energy_col, pose = i)
        } else if (fusion_method == "Median") {
          fused_data <- median_value(filtered_data, energy_col)
        } else if (fusion_method == "Arithmetic") {
          fused_data <- arithmetic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Geometric") {
          fused_data <- geometric_mean(filtered_data, energy_col)
        } else if (fusion_method == "Harmonic") {
          fused_data <- harmonic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Euclidean") {
          fused_data <- euclidean_mean(filtered_data, energy_col)
        }
        
        # Merge the calculated values back with experimental data
        merged_data <- left_join(fused_data, data_protein, by = "Title")

        # Detect the correct column name after merging
        fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

        # Fix .x and .y issues in merged_data
        if (paste0(fusion_col_name, ".x") %in% colnames(merged_data)) {
          merged_data <- merged_data %>%
            rename(!!fusion_col_name := !!sym(paste0(fusion_col_name, ".x"))) %>%
            select(-one_of(paste0(fusion_col_name, ".y")))
        }

        # Iterate over IC50 and pIC50 for correlation analysis
        for (exp_variable in exp_variables) {
          
          # Ensure there are no missing values in the correlation calculation
          valid_data <- merged_data %>%
            filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

          # Count the number of valid data points
          n_used <- nrow(valid_data)
          
          if (n_used < 3) {
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = NA,
              Spearman = NA,
              MAE = NA,
              RMSE = NA,
              IRM = NA,  # Nueva columna para IRM
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
            
            cat("Warning: No valid data for", fusion_method, "- Pose", i, "- EnergyTerm:", energy_col, "- ExpVariable:", exp_variable, "- Protein:", protein_name, "\n")
            
          } else {
            # Compute Pearson and Spearman correlations
            cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
            cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
            
            # Compute MAE, RMSE, and IRM
            mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
            rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
            irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  # Nuevo cálculo de IRM
            
            # Store results in dataframe
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = round(cor_pearson, 3),
              Spearman = round(cor_spearman, 3),
              MAE = round(mae_value, 3),
              RMSE = round(rmse_value, 3),
              IRM = round(irm_value, 3),  
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
          }
          
          # Append results to the list instead of `bind_rows()`
          results_list <- append(results_list, list(pose_df))
          
        }  # End iteration over IC50 and pIC50
      }  # End iteration over energy terms
    }  # End iteration over poses
    
    # Garbage collection every method to prevent memory overload
    gc()
  }  # End iteration over Data Fusion Methods
}  # End iteration over protein datasets

# Combine all results into a single dataframe
results_df <- do.call(rbind, results_list)

# Save results to an Excel file
write.xlsx(results_df, output_filename)

cat("Data successfully saved to", output_filename, "\n")
gc()

```

# **Variant 2 - 1 (IFD):**

### **Load data**

-   Modify `data_path` from **Load data**:

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/Data_Variant_2_1_IFD/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+") 

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)
    
    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)

```

### Correlation Data Analysis

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# Define the Data Fusion Methods
# -------------------------

apply_data_fusion <- function(data, method, energy_col, pose = NULL) {
  if (method == "Median") {
    return(median_value(data, energy_col))
  } else if (method == "Minimum") {
    return(minimum_value(data, energy_col, pose))  
  } else if (method == "Arithmetic") {
    return(arithmetic_mean(data, energy_col))
  } else if (method == "Geometric") {
    return(geometric_mean(data, energy_col))
  } else if (method == "Harmonic") {
    return(harmonic_mean(data, energy_col))
  } else if (method == "Euclidean") {
    return(euclidean_mean(data, energy_col))
  } else {
    stop("Invalid Data Fusion Method")
  }
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_2_1_IFD.xlsx"

# Define the variant and method
Variant <- "2_1"
Method <- "IFD"

# Define the energy-related columns to be analyzed
energy_columns <- c("docking.score", "glide.ligand.efficiency", 
                    "glide.gscore", "glide.emodel", "glide.energy", 
                    "glide.einternal")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# -------------------------
# Create a dataframe to store the results
# -------------------------

results_df <- data.frame(Protein = character(),
                         ExpVariable = character(),
                         DataFusion = character(),
                         EnergyTerm = character(),
                         Pose = integer(),
                         Pearson = numeric(),
                         Spearman = numeric(),
                         MAE = numeric(),
                         RMSE = numeric(),
                         IRM = numeric(),  # Nueva columna para IRM
                         N = integer(),
                         Variant = character(),
                         Method = character(),
                         stringsAsFactors = FALSE)

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]

  for (fusion_method in fusion_methods) {
    
    start_pose <- ifelse(fusion_method == "Minimum", 1, 2)

    for (i in start_pose:10) {
      
      # Filter data by poses
      filtered_data <- data_protein %>%
        filter(Poses <= i)  

      for (energy_col in energy_columns) {
        
        # Apply the Data Fusion technique
        if (fusion_method == "Minimum") {
          fused_data <- apply_data_fusion(filtered_data, fusion_method, energy_col, pose = i)  
        } else {
          fused_data <- apply_data_fusion(filtered_data, fusion_method, energy_col)
        }
        
        # Detect the correct name of the merged column
        fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

        # Merge with experimental data
        merged_data <- left_join(fused_data, data_protein[, c("Title", exp_variables)], by = "Title")
        
        # Iterate over IC50 and pIC50
        for (exp_variable in exp_variables) {
          
          # Filter non-null values for calculations
          valid_data <- merged_data %>%
            filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

          n_used <- nrow(valid_data)
          
          if (n_used < 3) {
            pose_df <- data.frame(
              Protein = protein_name,  
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = NA,
              Spearman = NA,
              MAE = NA,
              RMSE = NA,
              IRM = NA,  # We save NA in MRI if there is not enough data
              N = n_used,
              Variant = Variant,  
              Method = Method,  
              stringsAsFactors = FALSE
            )
            
          } else {
            # Calculating correlations and metrics
            cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
            cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
            mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
            rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
            irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  # Nuevo cálculo de IRM
            
            # Save results to the data frame
            pose_df <- data.frame(
              Protein = protein_name,  
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = round(cor_pearson, 3),
              Spearman = round(cor_spearman, 3),
              MAE = round(mae_value, 3),
              RMSE = round(rmse_value, 3),
              IRM = round(irm_value, 3),  # We added IRM
              N = n_used,
              Variant = Variant,  
              Method = Method,  
              stringsAsFactors = FALSE
            )
          }
          
          # Add results to the final dataframe
          results_df <- bind_rows(results_df, pose_df)
          
        }  # End iteration on IC50 and pIC50
      }  # End iteration on energy terms
    }  # End iteration over poses
  }  # End iteration on fusion methods
}  # End iteration on proteins

# Save results in Excel
write.xlsx(results_df, output_filename)

# Success message
cat("Data successfully saved to", output_filename, "\n")

```

# **Variant 2 - 2 (MMGBSA):**

### **Load data**

-   Modify `data_path` from **Load data**:

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/Data_Variant_2_2_MMGBSA/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+")

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)

    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists (it will be recalculated)
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)
```

### Correlation Data Analysis

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_2_2_MMGBSA_100poses.xlsx"

# Define the variant and method
Variant <- "2_2"
Method <- "MMGBSA"

# Define the energy-related columns to be analyzed
energy_columns <- c("MMGBSABind", "MMGBSABind_Coulomb", "MMGBSABind_Hbond", 
                    "MMGBSABind_Lipo", "MMGBSABind_Solv_GB", "MMGBSABind_vdW",
                    "Prime_MMGBSA_ligand_efficiency")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Create an empty list to store results before combining
# -------------------------

results_list <- list()  

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  
  cat("\n🔎 Processing Protein:", protein_name, "\n")
  
  # Iterate over each Data Fusion Method
  for (fusion_method in fusion_methods) {
    
    start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
    #max_poses <- ifelse(fusion_method == "Minimum", 10, 17)  # Define max poses
    max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"
    
    # Iterate over the number of poses
    for (i in start_pose:max_poses) {
      
      # Filter the dataset to include only the top `i` poses
      filtered_data <- data_protein %>%
        filter(Poses <= i)

      # Iterate over each energy-related variable
      for (energy_col in energy_columns) {
        
        # Apply the selected Data Fusion Method
        if (fusion_method == "Minimum") {
          fused_data <- minimum_value(filtered_data, energy_col, pose = i)
        } else if (fusion_method == "Median") {
          fused_data <- median_value(filtered_data, energy_col)
        } else if (fusion_method == "Arithmetic") {
          fused_data <- arithmetic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Geometric") {
          fused_data <- geometric_mean(filtered_data, energy_col)
        } else if (fusion_method == "Harmonic") {
          fused_data <- harmonic_mean(filtered_data, energy_col)
        } else if (fusion_method == "Euclidean") {
          fused_data <- euclidean_mean(filtered_data, energy_col)
        }
        
        # Merge the calculated values back with experimental data
        merged_data <- left_join(fused_data, data_protein, by = "Title")

        # Detect the correct column name after merging
        fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

        # Fix .x and .y issues in merged_data
        if (paste0(fusion_col_name, ".x") %in% colnames(merged_data)) {
          merged_data <- merged_data %>%
            rename(!!fusion_col_name := !!sym(paste0(fusion_col_name, ".x"))) %>%
            select(-one_of(paste0(fusion_col_name, ".y")))
        }

        # Iterate over IC50 and pIC50 for correlation analysis
        for (exp_variable in exp_variables) {
          
          # Ensure there are no missing values in the correlation calculation
          valid_data <- merged_data %>%
            filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

          # Count the number of valid data points
          n_used <- nrow(valid_data)
          
          if (n_used < 3) {
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = NA,
              Spearman = NA,
              MAE = NA,
              RMSE = NA,
              IRM = NA,  # Nueva columna para IRM
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
            
            cat(" Warning: No valid data for", fusion_method, "- Pose", i, "- EnergyTerm:", energy_col, "- ExpVariable:", exp_variable, "- Protein:", protein_name, "\n")
            
          } else {
            # Compute Pearson and Spearman correlations
            cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
            cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
            
            # Compute MAE, RMSE, and IRM
            mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
            rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
            irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  # ✅ Nuevo cálculo de IRM
            
            # Store results in dataframe
            pose_df <- data.frame(
              Protein = protein_name,
              ExpVariable = exp_variable,
              DataFusion = fusion_method,
              EnergyTerm = energy_col,
              Pose = i,
              Pearson = round(cor_pearson, 3),
              Spearman = round(cor_spearman, 3),
              MAE = round(mae_value, 3),
              RMSE = round(rmse_value, 3),
              IRM = round(irm_value, 3),  
              N = n_used,
              Variant = Variant,
              Method = Method,
              stringsAsFactors = FALSE
            )
          }
          
          # Append results to the list instead of `bind_rows()`
          results_list <- append(results_list, list(pose_df))
          
        }  # End iteration over IC50 and pIC50
      }  # End iteration over energy terms
    }  # End iteration over poses
    
    # Garbage collection every method to prevent memory overload
    gc()
  }  # End iteration over Data Fusion Methods
}  # End iteration over protein datasets

# Combine all results into a single dataframe
results_df <- do.call(rbind, results_list)

# Save results to an Excel file
write.xlsx(results_df, output_filename)

cat("Data successfully saved to", output_filename, "\n")
gc()

```

# **Variant 3 (ED)**

Need modify:

-   Modify `data_path` from **Load data**:

-   Modify `energy_columns` [docking or MM-GBSA terms] from **Correlation Data Analysis**

-   Modify `output_filename` from **Correlation Data Analysis**

-   Modify `Variant` and `Method` from **Correlation Data Analysis**

### **Load data**

Load data automatically from the data_path

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/02_Data_Variants/Data_Variant_3_ED/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+") 

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)
    
    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)

```

### Correlation Data Analysis

```{r}

# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export
library(stringr)   # For regex operations

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_3_ED.xlsx"

# Define the variant and method
Variant <- "3"
Method <- "ED"

# Define whether to compute Data Fusion across all frames or per frame
fusion_mode <- "both"  # Options: "combined", "per_frame", or "both"

# Define the energy-related columns to be analyzed
energy_columns <- c("docking.score", "glide.ligand.efficiency", 
                    "glide.gscore", "glide.emodel", "glide.energy", 
                    "glide.einternal")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Create an empty list to store results before combining
# -------------------------

results_list <- list()  

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  
  cat("\n🔎 Processing Protein:", protein_name, "\n")

  # Define how to treat frames
  frame_groups <- list()

  if (fusion_mode == "combined" || fusion_mode == "both") {
    frame_groups[["All_Frames"]] <- data_protein  
  }
  
  if (fusion_mode == "per_frame" || fusion_mode == "both") {
    unique_frames <- unique(data_protein$Frame)
    for (frame_id in unique_frames) {
      frame_groups[[as.character(frame_id)]] <- filter(data_protein, Frame == frame_id)
    }
  }

  # Iterate over each frame grouping option (all or separately)
  for (frame_id in names(frame_groups)) {
    frame_data <- frame_groups[[frame_id]]

    #Ensure that the `Frame` column is a string
    frame_data$Frame <- as.character(frame_data$Frame)

    # Iterate over each Data Fusion Method
    for (fusion_method in fusion_methods) {
      
      start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
      #max_poses <- ifelse(fusion_method == "Minimum", 10, 30)  # 
      max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"
      
      # Iterate over the number of poses
      for (i in start_pose:max_poses) {
        
        # Check if there are enough poses in the dataset
        available_poses <- unique(frame_data$Poses)
        if (!i %in% available_poses) {
          cat(" Skipping Pose", i, "for", fusion_method, "- Frame:", frame_id, "(Not enough poses available)\n")
          next  # Salta la iteración si la pose no existe
        }
        
        # Filter the dataset to include only the top `i` poses
        filtered_data <- frame_data %>%
          filter(Poses <= i)

        # Iterate over each energy-related variable
        for (energy_col in energy_columns) {
          
          # Apply the selected Data Fusion Method
          if (fusion_method == "Minimum") {
            fused_data <- minimum_value(filtered_data, energy_col, pose = i)
          } else if (fusion_method == "Median") {
            fused_data <- median_value(filtered_data, energy_col)
          } else if (fusion_method == "Arithmetic") {
            fused_data <- arithmetic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Geometric") {
            fused_data <- geometric_mean(filtered_data, energy_col)
          } else if (fusion_method == "Harmonic") {
            fused_data <- harmonic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Euclidean") {
            fused_data <- euclidean_mean(filtered_data, energy_col)
          }

          # Merge the calculated values back with experimental data
          merged_data <- suppressWarnings(
            left_join(fused_data, frame_data, by = "Title")
          )

          # Detect the correct column name after merging
          fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

          # Fix .x and .y issues in merged_data
          if (paste0(fusion_col_name, ".x") %in% colnames(merged_data)) {
            merged_data <- merged_data %>%
              rename(!!fusion_col_name := !!sym(paste0(fusion_col_name, ".x"))) %>%
              select(-one_of(paste0(fusion_col_name, ".y")))
          }

          # Iterate over IC50 and pIC50 for correlation analysis
          for (exp_variable in exp_variables) {
            
            # Ensure there are no missing values in the correlation calculation
            valid_data <- merged_data %>%
              filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

            # Count the number of valid data points
            n_used <- nrow(valid_data)
            
            if (n_used < 3) {
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id,  # Ahora almacena el frame correctamente (ej. "frame280")
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = NA,
                Spearman = NA,
                MAE = NA,
                RMSE = NA,
                IRM = NA,  # Nueva columna para IRM
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
              
              cat(" Warning: No valid data for", fusion_method, "- Pose", i, "- EnergyTerm:", energy_col, "- ExpVariable:", exp_variable, "- Protein:", protein_name, "- Frame:", frame_id, "\n")
              
            } else {
              # Compute Pearson and Spearman correlations
              cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
              cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
              
              # Compute MAE, RMSE, and IRM
              mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
              rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
              irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  # Nuevo cálculo de IRM
              
              # Store results in dataframe
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id,  # Ahora almacena el frame correctamente (ej. "frame280")
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = round(cor_pearson, 3),
                Spearman = round(cor_spearman, 3),
                MAE = round(mae_value, 3),
                RMSE = round(rmse_value, 3),
                IRM = round(irm_value, 3),  # Agregado IRM
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
            }
            
            # Append results to the list instead of `bind_rows()`
            results_list <- append(results_list, list(pose_df))
            
          }  # End iteration over IC50 and pIC50
        }  # End iteration over energy terms
      }  # End iteration over poses
      gc()  # Garbage collection
    }  # End iteration over frames
  }
}

# Combine all results into a single dataframe
results_df <- do.call(rbind, results_list)

# Save results to an Excel file
write.xlsx(results_df, output_filename)

cat("Data successfully saved to", output_filename, "\n")
gc()

```

# **Variant 4 (MM-GBSA)**

Need modify:

-   Modify `data_path` from **Load data**:

-   Modify `energy_columns` [docking or MM-GBSA terms] from **Correlation Data Analysis**

-   Modify `output_filename` from **Correlation Data Analysis**

-   Modify `Variant` and `Method` from **Correlation Data Analysis**

### **Load data**

Load data automatically from the data_path

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "/media/daniel/DATA2TB/DROP/Dropbox/UCM/Fondecyt_Iniciacion_2022/Aprovado/Etapas/Paper1/Paper_Hp/Analysis/Data_Variant_4_MMGBSA/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+")

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)

    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists (it will be recalculated)
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}
# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)

## if the result is zero, then change title by Title in xlsx files!
```

### Correlation Data Analysis

```{r}

# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_4_MMGBSA.xlsx"

# Define the variant and method
Variant <- "4"
Method <- "MMGBSA"

# Define whether to compute Data Fusion across all frames or per frame
fusion_mode <- "both"  # Options: "combined", "per_frame", or "both"

# Define the energy-related columns to be analyzed
energy_columns <- c("MMGBSABind", "MMGBSABind_Coulomb", "MMGBSABind_Hbond", 
                    "MMGBSABind_Lipo", "MMGBSABind_Solv_GB", "MMGBSABind_vdW",
                    "Prime_MMGBSA_ligand_efficiency")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Create an empty list to store results before combining
# -------------------------

results_list <- list()  

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  
  cat("\n🔎 Processing Protein:", protein_name, "\n")

  # Define how to treat frames
  frame_groups <- list()

  if (fusion_mode == "combined" || fusion_mode == "both") {
    frame_groups[["All_Frames"]] <- data_protein  
  }
  
  if (fusion_mode == "per_frame" || fusion_mode == "both") {
    unique_frames <- unique(data_protein$Frame)
    for (frame_id in unique_frames) {
      frame_groups[[as.character(frame_id)]] <- filter(data_protein, Frame == frame_id)
    }
  }

 
  for (frame_id in names(frame_groups)) {
    frame_data <- frame_groups[[frame_id]]

   
    frame_data$Frame <- as.character(frame_data$Frame)

    # Iterate over each Data Fusion Method
    for (fusion_method in fusion_methods) {
      
      start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
      #max_poses <- ifelse(fusion_method == "Minimum", 10, 10)  # Define max poses
      max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"

      # Iterate over the number of poses
      for (i in start_pose:max_poses) {
        
        # Filter the dataset to include only the top `i` poses
        filtered_data <- frame_data %>%
          filter(Poses <= i)

        # Iterate over each energy-related variable
        for (energy_col in energy_columns) {
          
          # Apply the selected Data Fusion Method
          if (fusion_method == "Minimum") {
            fused_data <- minimum_value(filtered_data, energy_col, pose = i)
          } else if (fusion_method == "Median") {
            fused_data <- median_value(filtered_data, energy_col)
          } else if (fusion_method == "Arithmetic") {
            fused_data <- arithmetic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Geometric") {
            fused_data <- geometric_mean(filtered_data, energy_col)
          } else if (fusion_method == "Harmonic") {
            fused_data <- harmonic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Euclidean") {
            fused_data <- euclidean_mean(filtered_data, energy_col)
          }

          # Merge the calculated values back with experimental data
          merged_data <- suppressWarnings(left_join(fused_data, frame_data, by = "Title"))

          # Detect the correct column name after merging
          fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

          # Fix .x and .y issues in merged_data
          if (paste0(fusion_col_name, ".x") %in% colnames(merged_data)) {
            merged_data <- merged_data %>%
              rename(!!fusion_col_name := !!sym(paste0(fusion_col_name, ".x"))) %>%
              select(-one_of(paste0(fusion_col_name, ".y")))
          }

          # Iterate over IC50 and pIC50 for correlation analysis
          for (exp_variable in exp_variables) {
            
            # Ensure there are no missing values in the correlation calculation
            valid_data <- merged_data %>%
              filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

            # Count the number of valid data points
            n_used <- nrow(valid_data)
            
            if (n_used < 3) {
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id,  
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = NA,
                Spearman = NA,
                MAE = NA,
                RMSE = NA,
                IRM = NA,  
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
              
              cat(" Warning: No valid data for", fusion_method, "- Pose", i, "- EnergyTerm:", energy_col, "- ExpVariable:", exp_variable, "- Protein:", protein_name, "- Frame:", frame_id, "\n")
              
            } else {
              # Compute Pearson and Spearman correlations
              cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
              cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
              
              # Compute MAE, RMSE, and IRM
              mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
              rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
              irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])   
              
              # Store results in dataframe
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id,  
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = round(cor_pearson, 3),
                Spearman = round(cor_spearman, 3),
                MAE = round(mae_value, 3),
                RMSE = round(rmse_value, 3),
                IRM = round(irm_value, 3), 
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
            }
            
            # Append results to the list instead of `bind_rows()`
            results_list <- append(results_list, list(pose_df))
            
          }  # End iteration over IC50 and pIC50
        }  # End iteration over energy terms
      }  # End iteration over poses
      
      # Garbage collection every 10 poses to prevent memory overload
      gc()
    }  # End iteration over frames
  }  # End iteration over Data Fusion Methods
}  # End iteration over pr

# Combine all results into a single dataframe
results_df <- do.call(rbind, results_list)

# Save results to an Excel file
write.xlsx(results_df, output_filename)

cat("Data successfully saved to", output_filename, "\n")
gc()
```

# **Variant 5 (QPLD):**

### **Load data**

-   Modify `data_path` from **Load data**:

```{r}
# Load required libraries
library(openxlsx)
library(dplyr)
library(stringr)

# -------------------------
# User Configuration
# -------------------------

# Define the directory containing the Excel files
data_path <- "02_Data_Variants/Data_Variant_5_QPLD/"

# List all .xlsx files in the directory
file_list <- list.files(path = data_path, pattern = "*.xlsx", full.names = TRUE)

# Extract protein names from filenames (Assumes filenames are formatted as: combined_data_Protein.xlsx)
protein_names <- str_extract(basename(file_list), "(?<=combined_data_)[^\\.]+")

# -------------------------
# Load Data
# -------------------------

# Read each Excel file into a list of data frames
data_list <- lapply(seq_along(file_list), function(i) {
  file <- file_list[i]
  protein <- protein_names[i]
  
  # Read file if it exists
  if (file.exists(file)) {
    df <- read.xlsx(file, sheet = 1)
    
    # Remove pIC50.Exp if it exists
    if ("pIC50.Exp" %in% colnames(df)) {
      df <- df %>% select(-pIC50.Exp)
    }
    
    # Remove pIC50 if it exists
    if ("pIC50" %in% colnames(df)) {
      df <- df %>% select(-pIC50)
    }
    
    # Remove last "_X" from Title
    df <- df %>% mutate(Title = str_replace(Title, "_\\d+$", ""))
    
    # Calculate pIC50 from IC50 (IC50 is in µM, needs to be converted to M)
    df <- df %>% mutate(pIC50 = -log10(IC50 * 1e-6))
    
    # Add protein name as a new column
    df <- df %>% mutate(Protein = protein)
    
    return(df)
  } else {
    warning(paste("File not found:", file))
    return(NULL)
  }
})

# Assign names to the list for easier reference
names(data_list) <- protein_names

# -------------------------
# Filter Data
# -------------------------

# Function to filter data
filter_data <- function(df) {
  df %>% 
    filter(IC50 != 800 & !is.na(IC50) & pIC50 != 3.09 & !is.na(pIC50))
}

# Apply filtering to each dataset
filtered_data_list <- lapply(data_list, function(df) {
  if (!is.null(df)) {
    filter_data(df)
  } else {
    NULL
  }
})

# -------------------------
# Summary of Removed Values
# -------------------------

# Count missing values and occurrences of 800 in IC50 and 3.09 in pIC50 before filtering
missing_summary <- data.frame(
  Protein = protein_names,
  Missing_IC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$IC50)) else NA),
  IC50_800 = sapply(data_list, function(df) if (!is.null(df)) sum(df$IC50 == 800, na.rm = TRUE) else NA),
  Missing_pIC50 = sapply(data_list, function(df) if (!is.null(df)) sum(is.na(df$pIC50)) else NA),
  pIC50_309 = sapply(data_list, function(df) if (!is.null(df)) sum(df$pIC50 == 3.09, na.rm = TRUE) else NA)
)

# Display the summary of removed values
print(missing_summary)

# The `filtered_data_list` now contains cleaned datasets for each protein.

```

### **Verify the number of compounds**

Prints the number of unique compounds for each file

```{r}

# Function to count unique compounds (Title)
count_unique_titles <- function(df) {
  if (!is.null(df)) {
    return(length(unique(df$Title)))
  } else {
    return(NA)
  }
}

# Apply function to each dataset in the list
unique_counts <- sapply(filtered_data_list, count_unique_titles)

# Display the results
print(unique_counts)
```

### Correlation Data Analysis

```{r}
# Load necessary libraries
library(dplyr)
library(openxlsx)  # Required for Excel export
library(stringr)   # For regex operations

# -------------------------
# Define IRM Function
# -------------------------

# IRM: Inlier Ratio Metric - Proporción de predicciones dentro del umbral de error
calculate_irm <- function(actual, predicted, threshold = 1.0) {
  mean(abs(actual - predicted) <= threshold, na.rm = TRUE)  
}

# -------------------------
# User Configuration
# -------------------------

# Define the output filename
output_filename <- "Correlation_Variant_5_QPLD_1E9Y.xlsx"

# Define the variant and method
Variant <- "5"
Method <- "QPLD"

# Define whether to compute Data Fusion across all frames or per frame
fusion_mode <- "both"  # Options: "combined", "per_frame", or "both"

# Define the energy-related columns to be analyzed
energy_columns <- c("docking.score", "glide.ligand.efficiency", 
                    "glide.gscore", "glide.emodel", "glide.energy", 
                    "glide.einternal")

# Define experimental variables to compare (IC50 and pIC50)
exp_variables <- c("IC50", "pIC50")

# Define the list of Data Fusion Methods to apply
fusion_methods <- c("Median", "Minimum", "Arithmetic", "Geometric", "Harmonic", "Euclidean")

# -------------------------
# Create an empty list to store results before combining
# -------------------------

results_list <- list()  # Usar lista en vez de `bind_rows()` para eficiencia

# -------------------------
# Iterate over each protein dataset in `filtered_data_list`
# -------------------------

for (protein_name in names(filtered_data_list)) {
  data_protein <- filtered_data_list[[protein_name]]
  
  cat("\n🔎 Processing Protein:", protein_name, "\n")

  # Definir cómo tratar los frames
  frame_groups <- list()

  if (fusion_mode == "combined" || fusion_mode == "both") {
    frame_groups[["All_Frames"]] <- data_protein  
  }
  
  if (fusion_mode == "per_frame" || fusion_mode == "both") {
    unique_frames <- unique(data_protein$Frame)
    for (frame_id in unique_frames) {
      frame_groups[[as.character(frame_id)]] <- filter(data_protein, Frame == frame_id)
    }
  }

  
  for (frame_id in names(frame_groups)) {
    frame_data <- frame_groups[[frame_id]]

    
    frame_data$Frame <- as.character(frame_data$Frame)

    # Iterate over each Data Fusion Method
    for (fusion_method in fusion_methods) {
      
      start_pose <- ifelse(fusion_method == "Minimum", 1, 2)
      #max_poses <- ifelse(fusion_method == "Minimum", 10, 100)  # Define max poses
      max_poses <- 100  # Set max poses to 100 for all methods, including "Minimum"
      
      # Iterate over the number of poses
      for (i in start_pose:max_poses) {
        
        # Filter the dataset to include only the top `i` poses
        filtered_data <- frame_data %>%
          filter(Poses <= i)

        # Iterate over each energy-related variable
        for (energy_col in energy_columns) {
          
          # Apply the selected Data Fusion Method
          if (fusion_method == "Minimum") {
            fused_data <- minimum_value(filtered_data, energy_col, pose = i)
          } else if (fusion_method == "Median") {
            fused_data <- median_value(filtered_data, energy_col)
          } else if (fusion_method == "Arithmetic") {
            fused_data <- arithmetic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Geometric") {
            fused_data <- geometric_mean(filtered_data, energy_col)
          } else if (fusion_method == "Harmonic") {
            fused_data <- harmonic_mean(filtered_data, energy_col)
          } else if (fusion_method == "Euclidean") {
            fused_data <- euclidean_mean(filtered_data, energy_col)
          }

          # Merge the calculated values back with experimental data
          merged_data <- suppressWarnings(
            left_join(fused_data, frame_data, by = "Title")
          )

          # Detect the correct column name after merging
          fusion_col_name <- colnames(fused_data)[which(colnames(fused_data) != "Title")]

          # Fix .x and .y issues in merged_data
          if (paste0(fusion_col_name, ".x") %in% colnames(merged_data)) {
            merged_data <- merged_data %>%
              rename(!!fusion_col_name := !!sym(paste0(fusion_col_name, ".x"))) %>%
              select(-one_of(paste0(fusion_col_name, ".y")))
          }

          # Iterate over IC50 and pIC50 for correlation analysis
          for (exp_variable in exp_variables) {
            
            # Ensure there are no missing values in the correlation calculation
            valid_data <- merged_data %>%
              filter(!is.na(.data[[exp_variable]]) & !is.na(.data[[fusion_col_name]]))

            # Count the number of valid data points
            n_used <- nrow(valid_data)
            
            if (n_used < 3) {
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id,  
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = NA,
                Spearman = NA,
                MAE = NA,
                RMSE = NA,
                IRM = NA,  
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
              
              cat("️ Warning: No valid data for", fusion_method, "- Pose", i, "- EnergyTerm:", energy_col, "- ExpVariable:", exp_variable, "- Protein:", protein_name, "- Frame:", frame_id, "\n")
              
            } else {
              # Compute Pearson and Spearman correlations
              cor_pearson <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs")
              cor_spearman <- cor(valid_data[[exp_variable]], valid_data[[fusion_col_name]], use = "complete.obs", method = "spearman")
              
              # Compute MAE, RMSE, and IRM
              mae_value <- mean(abs(valid_data[[exp_variable]] - valid_data[[fusion_col_name]]), na.rm = TRUE)
              rmse_value <- sqrt(mean((valid_data[[exp_variable]] - valid_data[[fusion_col_name]])^2, na.rm = TRUE))
              irm_value <- calculate_irm(valid_data[[exp_variable]], valid_data[[fusion_col_name]])  
              
              # Store results in dataframe
              pose_df <- data.frame(
                Protein = protein_name,
                Frame = frame_id, 
                ExpVariable = exp_variable,
                DataFusion = fusion_method,
                EnergyTerm = energy_col,
                Pose = i,
                Pearson = round(cor_pearson, 3),
                Spearman = round(cor_spearman, 3),
                MAE = round(mae_value, 3),
                RMSE = round(rmse_value, 3),
                IRM = round(irm_value, 3),  
                N = n_used,
                Variant = Variant,
                Method = Method,
                stringsAsFactors = FALSE
              )
            }
            
            # Append results to the list instead of `bind_rows()`
            results_list <- append(results_list, list(pose_df))
            
          }  # End iteration over IC50 and pIC50
        }  # End iteration over energy terms
      }  # End iteration over poses
      
      # Garbage collection every 10 poses to prevent memory overload
      gc()
    }  # End iteration over frames
  }  # End iteration over Data Fusion Methods
}  # End iteration over protein datasets

#  Combine all results into a single dataframe
results_df <- do.call(rbind, results_list)

# Save results to an Excel file
write.xlsx(results_df, output_filename)

cat("✅ Data successfully saved to", output_filename, "\n")
gc()

```
