---
title: "Comparative Data Analysis of Virtual Screening
Methodologies for Predicting Urease Inhibitory Activity"
output: html_notebook
---

# **Load correlation data & merging**

```{r}

# Load necessary libraries
library(dplyr)
library(readxl)
library(openxlsx)
library(stringr)  # Necessary to extract the first number of NVariant

# -------------------------
# Define File Paths
# -------------------------

# Base path where the files are located
base_path <- "03_Correlations/"

# List of correlation files
file_list <- list(
  paste0(base_path, "Correlation_Variant_1_1_Docking.xlsx"),
  paste0(base_path, "Correlation_Variant_1_2_QPLD_100poses.xlsx"),
  paste0(base_path, "Correlation_Variant_1_3_MMGBSA_100poses.xlsx"),
  paste0(base_path, "Correlation_Variant_2_1_IFD.xlsx"),
  paste0(base_path, "Correlation_Variant_2_2_MMGBSA_100poses.xlsx"),
  paste0(base_path, "Correlation_Variant_3_ED.xlsx"),
  paste0(base_path, "Correlation_Variant_4_MMGBSA.xlsx"),
  paste0(base_path, "Correlation_Variant_5_QPLD_1E9Y.xlsx"),
  paste0(base_path, "Correlation_Variant_5_QPLD_6QSU.xlsx"),
  paste0(base_path, "Correlation_Variant_5_QPLD_6ZJA.xlsx")
)

# -------------------------
# Function to Process Files
# -------------------------

process_correlation_file <- function(file) {
  cat("\n📂 Processing:", file, "\n")
  
  # Efficiently read Excel file
  df <- read_excel(file, col_types = "guess")
  
  # Rename Variant to NVariant and create new Variant column
  df <- df %>%
    rename(NVariant = Variant) %>%
    mutate(
      # Extract the first number from NVariant and convert to integer
      Variant = as.integer(str_extract(NVariant, "^[0-9]+"))
    )
  
  # Add Frame column for Variants 1 and 2 (set to "0" as character)
  if (grepl("Variant_1_|Variant_2_", file)) {
    df <- df %>%
      mutate(Frame = "0") %>%
      relocate(Frame, .after = Protein)  # Ensure Frame is in second position
  } else {
    # Ensure Frame exists and is a character type in all files
    df <- df %>%
      mutate(Frame = as.character(Frame))
  }
  
  return(df)
}

# -------------------------
# Load and Process All Files
# -------------------------

correlation_data_list <- lapply(file_list, process_correlation_file)

# -------------------------
# Merge All Dataframes
# -------------------------

# Bind all processed dataframes together
combined_correlation_data <- bind_rows(correlation_data_list)

# Save path for the combined file in RDS
rds_file <- paste0(base_path, "Merged_Correlation_Data.rds")

# Save in RDS format (much more efficient than Excel)
saveRDS(combined_correlation_data, rds_file)

# Optionally, we also save in Excel
#xlsx_file <- paste0(base_path, "Merged_Correlation_Data.xlsx")
#write.xlsx(combined_correlation_data, xlsx_file)

cat("\n✅ All files processed and merged successfully!\n")
cat("📂 Saved as:\n - RDS:", rds_file)


```

# **Load rds file with all data**

This is

```{r}
# Load dataframe in R from RDS
df <- readRDS("03_Correlations/Merged_Correlation_Data.rds")

# Remove rows with any NA values
df <- na.omit(df)

# Check the dimensions after removal
dim(df)

original_n <- nrow(readRDS("03_Correlations/Merged_Correlation_Data.rds"))
new_n <- nrow(df)

cat("Rows removed due to NA values:", original_n - new_n, "\n")
```

# **Is Considering More Poses Better?**

## Fig. 2

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(gridExtra)

# Function to compute summary statistics
compute_summary <- function(data) {
  data %>%
    group_by(DataFusion) %>%
    summarise(
      Mean_Pearson = mean(Pearson, na.rm = TRUE),
      Mean_Spearman = mean(Spearman, na.rm = TRUE),
      Mean_MAE = mean(MAE, na.rm = TRUE),
      Mean_RMSE = mean(RMSE, na.rm = TRUE),
      Mean_IRM = mean(IRM, na.rm = TRUE),
      N = n()
    ) %>%
    arrange(desc(Mean_Spearman))  # Sorting by highest Spearman correlation
}

# 🔹 Filter only IC50 data
df_ic50 <- df %>%
  filter(ExpVariable == "IC50")

# 🔹 Compute summary statistics for Poses 1-10
summary_10_poses <- df_ic50 %>%
  filter(Pose >= 1, Pose <= 10) %>%
  compute_summary() %>%
  rename_with(~ paste0(., "_10Poses"))

# 🔹 Compute summary statistics for Poses 1-100
summary_100_poses <- df_ic50 %>%
  filter(Pose >= 1, Pose <= 100) %>%
  compute_summary() %>%
  rename_with(~ paste0(., "_100Poses"))

# 🔹 Standardize column names before joining
summary_10_poses <- summary_10_poses %>%
  rename(DataFusion = DataFusion_10Poses)

summary_100_poses <- summary_100_poses %>%
  rename(DataFusion = DataFusion_100Poses)

# 🔹 Find common Data Fusion methods
common_datafusion <- intersect(summary_10_poses$DataFusion, summary_100_poses$DataFusion)

# 🔹 Filter both datasets to retain only common DataFusion methods
summary_10_poses_filtered <- summary_10_poses %>%
  filter(DataFusion %in% common_datafusion)

summary_100_poses_filtered <- summary_100_poses %>%
  filter(DataFusion %in% common_datafusion)

# 🔹 Merge the two datasets
comparison_df <- left_join(summary_10_poses_filtered, summary_100_poses_filtered, by = "DataFusion")

# 🔹 Compute relative change between 10 and 100 poses for each metric
comparison_df <- comparison_df %>%
  mutate(
    Pearson_Change = (Mean_Pearson_100Poses - Mean_Pearson_10Poses) / abs(Mean_Pearson_10Poses) * 100,
    Spearman_Change = (Mean_Spearman_100Poses - Mean_Spearman_10Poses) / abs(Mean_Spearman_10Poses) * 100,
    MAE_Change = (Mean_MAE_100Poses - Mean_MAE_10Poses) / abs(Mean_MAE_10Poses) * 100,
    RMSE_Change = (Mean_RMSE_100Poses - Mean_RMSE_10Poses) / abs(Mean_RMSE_10Poses) * 100,
    IRM_Change = (Mean_IRM_100Poses - Mean_IRM_10Poses) / abs(Mean_IRM_10Poses) * 100
  )


# Define custom colors for Data Fusion Methods
fusion_colors <- c(
  "Median" = "#1f77b4",
  "Minimum" = "#ff7f0e",
  "Arithmetic" = "#2ca02c",
  "Geometric" = "#d62728",
  "Harmonic" = "#9467bd",
  "Euclidean" = "#8c564b"
)

# Define custom themes
custom_theme_left <- theme(
  legend.position = "none",
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text = element_text(colour = "black", face = "bold", size = 17),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  plot.title = element_text(size = 18, face = "bold")
)

custom_theme_right <- theme(
  legend.position = "none",
  axis.title.y = element_blank(),  # Remove Y-axis title
  axis.text.y = element_blank(),   # Remove Y-axis labels
  axis.ticks.y = element_blank(),  # Remove Y-axis ticks
  axis.title.x = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 17),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  plot.title = element_text(size = 18, face = "bold")
)

legend_theme <- theme(
  legend.position = "top",
  legend.title = element_blank(),
  legend.text = element_text(color = "black", size = 15)
)

# Generate individual plots
plot_pearson <- ggplot(comparison_df, aes(x = DataFusion, y = Pearson_Change, fill = DataFusion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fusion_colors) +
  scale_y_continuous(limits = c(-100, 180), breaks = seq(-100, 180, by = 40)) +
  coord_flip() +
  labs(title = "Pearson Correlation Change", x = "Data Fusion", y = "Change (%)") +
  custom_theme_right  # Remove Y-axis from second column

plot_spearman <- ggplot(comparison_df, aes(x = DataFusion, y = Spearman_Change, fill = DataFusion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fusion_colors) +
  scale_y_continuous(limits = c(-30, 10), breaks = seq(-30, 10, by = 10)) +
  coord_flip() +
  labs(title = "Spearman Correlation Change", x = "Data Fusion", y = "Change (%)") +
  custom_theme_left
 
plot_rmse <- ggplot(comparison_df, aes(x = DataFusion, y = RMSE_Change, fill = DataFusion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fusion_colors) +
  coord_flip() +
  labs(title = "RMSE Change", x = "Data Fusion", y = "Change (%)") +
  custom_theme_right

plot_irm <- ggplot(comparison_df, aes(x = DataFusion, y = IRM_Change, fill = DataFusion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fusion_colors) +
  scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 10)) +
  coord_flip() +
  labs(title = "IRM Change", x = "Data Fusion", y = "Change (%)") +
  custom_theme_left

plot_mae <- ggplot(comparison_df, aes(x = DataFusion, y = MAE_Change, fill = DataFusion)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = fusion_colors) +
  coord_flip() +
  labs(title = "MAE Change", x = "Data Fusion", y = "Change (%)") +
  custom_theme_left

# Create a legend-only plot (completely blank except for the legend)
legend_plot <- ggplot(comparison_df, aes(x = DataFusion, fill = DataFusion)) +
  geom_bar(alpha = 10) +  # Make the bars invisible
  scale_fill_manual(values = fusion_colors) +
  labs(title = "") +
  theme_void() +  # Remove all background, shafts and grid
  theme(
    legend.position = "top",
    legend.text = element_text(size = 14),  # Increasing the size of the caption text
    legend.key.size = unit(1, "cm")  # Increase the size of colored boxes in the legend
  )

# Arrange all plots into a single panel using gridExtra
final_plot <- grid.arrange(
  plot_spearman, plot_pearson,
  plot_mae, plot_rmse,
  plot_irm, legend_plot,  # Legend in the last slot
  ncol = 2
)

# Save the updated figure

ggsave("04_Figures/R/Fig2.tiff", plot = final_plot, width = 12, height = 14, dpi = 300)

# Export comparison_df to Excel
library(openxlsx)
write.xlsx(comparison_df, "Comparison_DataFusion_10vs100Poses.xlsx")

```

# **How do Data Fusion methods behave across different proteins?**

### Complete (Fig. 3A)

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(formattable)
library(webshot)
library(htmlwidgets)


# -------------------------
# Filter only IC50 data and limit to poses ≤ 10
# -------------------------
df_ic50 <- df %>%
  filter(ExpVariable == "IC50", Pose <= 10)  # Filter Pose ≤ 10

# -------------------------
# Compute Summary Statistics per Data Fusion method (Spearman only)
# -------------------------
summary_df <- df_ic50 %>%
  group_by(DataFusion) %>%
  summarise(
    N = n(),
    Mean = mean(Spearman, na.rm = TRUE),
    SD = sd(Spearman, na.rm = TRUE),
    Q1 = quantile(Spearman, 0.25, na.rm = TRUE),
    Median = median(Spearman, na.rm = TRUE),
    Q3 = quantile(Spearman, 0.75, na.rm = TRUE),
    # Count of outliers above Q3
   "Outliers ≥ Q3" = sum(Spearman > quantile(Spearman, 0.75, na.rm = TRUE), na.rm = TRUE),
    # Count of outliers above or equal to 0.9
   "Outliers ≥ 0.9" = sum(Spearman >= 0.9, na.rm = TRUE),
    # Count of outliers exactly equal to 1
    "Outliers = 1.0" = sum(Spearman == 1, na.rm = TRUE)
  ) %>%
  arrange(desc(Mean))  # Sorting by highest Mean Spearman correlation


# Ensure PhantomJS is installed for image export
webshot::install_phantomjs()

# ------------------------- 
# Define Advanced Heatmap Gradient Function
# -------------------------
heatmap_gradient <- function(x) {
  normalize <- (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  colorRampPalette(c("white", "yellow", "orange", "red"))(100)[as.numeric(cut(normalize, breaks = 100))]
}

# Format summary_df to 3 decimal places
summary_df <- summary_df %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

# -------------------------
# Apply Heatmap Formatting to the Table
# -------------------------
formatted_table <- formattable(summary_df, list(
  Mean = color_tile("white", "#FF3D3D"),
  SD = color_tile("white", "#FF3D3D"),
  Q1 = color_tile("white", "#FF3D3D"),
  Median = color_tile("white", "#FF3D3D"),
  Q3 = color_tile("white", "#FF3D3D"),
  N = color_tile("white", "#FF3D3D"),
  "Outliers ≥ Q3" = color_tile("white", "#FF3D3D"),
  "Outliers ≥ 0.9" = color_tile("white", "#FF3D3D"),
  "Outliers = 1.0" = color_tile("white", "#FF3D3D")
))

# -------------------------
# Save the Table as an HTML File
# -------------------------
html_file <- "Fig3A.html"
saveWidget(as.htmlwidget(formatted_table), file = html_file, selfcontained = TRUE)

# -------------------------
# Convert HTML to PNG (Ensure PhantomJS is installed)
# -------------------------
png_file <- "Fig3A.png"
webshot(html_file, file = png_file, vwidth = 800, vheight = 200)  # Adjusted dimensions


```

### Per protein (Fig. 3B)

```{r}

# Load necessary libraries
library(dplyr)
library(formattable)
library(webshot)
library(htmlwidgets)

# Ensure PhantomJS is installed for image export
webshot::install_phantomjs()

# -------------------------
# Filter only IC50 data and limit to poses ≤ 10
# -------------------------
df_ic50 <- df %>%
  filter(ExpVariable == "IC50", Pose <= 10)  # Filter Pose ≤ 10

# -------------------------
# Compute Summary Statistics per Data Fusion method and Protein
# -------------------------
summary_df <- df_ic50 %>%
  group_by(Protein, DataFusion) %>%
  summarise(
    N = n(),
    Mean = mean(Spearman, na.rm = TRUE),
    SD = sd(Spearman, na.rm = TRUE),
    Q1 = quantile(Spearman, 0.25, na.rm = TRUE),
    Median = median(Spearman, na.rm = TRUE),
    Q3 = quantile(Spearman, 0.75, na.rm = TRUE),
    
    "Outliers ≥ Q3" = sum(Spearman > quantile(Spearman, 0.75, na.rm = TRUE), na.rm = TRUE),
    "Outliers ≥ 0.9" = sum(Spearman >= 0.9, na.rm = TRUE),
    "Outliers = 1.0" = sum(Spearman == 1, na.rm = TRUE)
  ) %>%
  arrange(Protein, DataFusion)  # Sorting by Protein first, then DataFusion

# -------------------------
# Round numeric values to 3 decimal places
# -------------------------
summary_df <- summary_df %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

# -------------------------
# Define Advanced Heatmap Gradient Function
# -------------------------
heatmap_gradient <- function(x) {
  normalize <- (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
  colorRampPalette(c("white", "yellow", "orange", "#FF6B6B"))(100)[as.numeric(cut(normalize, breaks = 100))]
}

# -------------------------
# Apply Heatmap Formatting
# -------------------------
formatted_table <- formattable(summary_df, list(
  Mean = color_tile("white", "#FF6B6B"),
  SD = color_tile("white", "#FF6B6B"),
  Q1 = color_tile("white", "#FF6B6B"),
  Median = color_tile("white", "#FF6B6B"),
  Q3 = color_tile("white", "#FF6B6B"),
  N = color_tile("white", "#FF6B6B"),
  "Outliers ≥ Q3" = color_tile("white", "#FF6B6B"),
  "Outliers ≥ 0.9" = color_tile("white", "#FF6B6B"),
  "Outliers = 1.0" = color_tile("white", "#FF6B6B")
))

# -------------------------
# Save Table as an Image
# -------------------------
# Save as an HTML file first
html_file <- "Fig3B.html"
saveWidget(as.htmlwidget(formatted_table), file = html_file, selfcontained = TRUE)

# Convert HTML to PNG
png_file <- "Fig3B.png"
#webshot(html_file, file = png_file, vwidth = 1800, vheight = 1000)
webshot(html_file, file = png_file, vwidth = 800, vheight = 200)

```

# Pose aggregation and data fusion: identifying optimal strategies across proteins

## Fig. 4A-D

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.title = element_text(size = 22, face = "bold"), # Adjust size as needed
  legend.text = element_text(color = "black", size = 18),
  axis.title = element_text(color = "black", face = "bold", size = 24),
  axis.text.x = element_text(colour = "black", face = "bold", size = 20, angle = 20, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 20),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")
)

# Filter only IC50 data with Spearman > 0.25
df_ic50 <- df %>%
  filter(ExpVariable == "IC50",  Pose <= 100, Pose > -1, Spearman > 0 )

# Summarize Spearman correlation per Data Fusion, Pose, and Protein
heatmap_data <- df_ic50 %>%
  group_by(Protein, DataFusion, Pose) %>%
  summarise(Mean_Spearman = max(Spearman, na.rm = TRUE), .groups = "drop")

# Generate the heatmap per Protein with Pose labels in intervals of 10
heatmap_DF <- ggplot(heatmap_data, aes(y = factor(Pose), x = DataFusion, fill = Mean_Spearman)) +
  geom_tile(color = "white") +  # Heatmap tiles with white border
  scale_fill_gradient2(low = "#33608C", mid = "white", high = "#B81840", midpoint = 0.5, na.value = "black", limits = c(0, 1)) +
  scale_y_discrete(breaks = seq(min(heatmap_data$Pose), max(heatmap_data$Pose), by = 9)) + # Show Poses in steps of 10
  labs(title = "",
       y = "Pose",
       x = "Data Fusion Method",
       fill = "Spearman correlation") +
  theme_minimal() +
  custom_theme +  # Apply custom theme
  facet_wrap(~ Protein, scales = "fixed")  # Separate heatmap per protein

heatmap_DF

ggsave("04_Figures/R/Fig4AD.tiff", plot = heatmap_DF, width = 12, height = 14, dpi = 300)
```

#    

## Fig. 4E-H

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.title = element_text(size = 22, face = "bold"), # Adjust size as needed
  legend.text = element_text(color = "black", size = 18),
  axis.title = element_text(color = "black", face = "bold", size = 24),
  axis.text.x = element_text(colour = "black", face = "bold", size = 20, angle = 20, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 20),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")
)

# Filter only IC50 data with Spearman > 0.3
df_ic50 <- df %>%  
   filter(ExpVariable == "IC50",  Pose <= 100, Pose > -1, Spearman > 0 )

# Summarize N per Data Fusion, Pose, and Protein
heatmap_data <- df_ic50 %>%
  group_by(Protein, DataFusion, Pose) %>%
  summarise(N = n(), .groups = "drop")  # Count the number of values

# Generate the heatmap per Protein
heatmap_N <- ggplot(heatmap_data, aes(y = factor(Pose), x = DataFusion, fill = N)) +
  geom_tile(color = "white") +  # Heatmap tiles with white border
  scale_fill_gradient(low = "#FEFFD9", high = "#146C36", na.value = "black", limits = c(0, 400)) +  # Custom color scale for N
  scale_y_discrete(breaks = seq(min(heatmap_data$Pose), max(heatmap_data$Pose), by = 9)) +  # Show Poses in steps of 10
  labs(title = "",
       y = "Pose",
       x = "Data Fusion Method",
       fill = "Number of data") +  # Adjust legend title
  theme_minimal() +
  custom_theme +  # Apply custom theme
  facet_wrap(~ Protein, scales = "fixed")  # Separate heatmap per protein

# Display plot
heatmap_N

# Save the figure
ggsave("04_Figures/R/Fig4EH.tiff", plot = heatmap_N, width = 12, height = 14, dpi = 300)
```

# Energy Terms

## Fig. 5

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(color = "black", size = 18),
  legend.title = element_text(size = 18, face = "bold"),  # Increased legend title size
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 16, angle = 20, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 16),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")  # Increases Protein label size
)

# Filter only IC50 data
df_ic50 <- df %>%  
  filter(ExpVariable == "IC50", Pose < 10 )  

# Summarize Spearman correlation per EnergyTerm, DataFusion, Variant, and Protein
heatmap_data <- df_ic50 %>%
  group_by(Protein, Variant, EnergyTerm, DataFusion) %>%
  summarise(Mean_Spearman = max(Spearman, na.rm = TRUE), .groups = "drop")

# Generate the heatmap faceted by Protein
heatmap_plot <- ggplot(heatmap_data, aes(x = DataFusion, y = EnergyTerm, fill = Mean_Spearman)) +
  geom_tile(color = "white") +  # Heatmap tiles with white border
  scale_fill_gradient2(low = "#53A054", mid = "white", high = "#540046", midpoint = 0) +  
  labs(title = "",  # No title for clean formatting
       x = "Data Fusion Method",
       y = "Energy Term",
       fill = "Spearman correlation") +
  theme_minimal() +
  custom_theme +  # Apply custom theme
  facet_wrap(~ Protein, scales = "fixed")  # Separate heatmap per Protein

# Display the plot
heatmap_plot

# Save the heatmap as TIFF
ggsave("04_Figures/R/Fig5.tiff", plot = heatmap_plot, width = 14, height = 12, dpi = 300)
```

# Methods comparison is VS: performance across metrics

## Fig. 6

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(stringr)
library(gridExtra)  # Alternative to patchwork for combining plots

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(color = "black", size = 18),
  legend.title = element_text(size = 18, face = "bold"),  # Increased legend title size
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 16, angle = 0, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 16),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")  # Increases Protein label size
)

# -------------------------
# Filter Data (IC50, Consider Methods, DataFusion = Minimum)
# -------------------------
df_ic50 <- df %>%
  filter(
    ExpVariable == "IC50",
    EnergyTerm %in% c("MMGBSABind_Coulomb", "docking.score"),  
    Pose <= 19,  
    DataFusion == "Minimum"  
  )

# -------------------------
# Summarize Maximum Metrics per Protein and Method
# -------------------------
heatmap_data <- df_ic50 %>%
  group_by(Protein, Method) %>%
  summarise(
    Spearman = max(Spearman, na.rm = TRUE),
    Pearson = max(Pearson, na.rm = TRUE),
    MAE = max(MAE, na.rm = TRUE),
    RMSE = max(RMSE, na.rm = TRUE),
    IRM = max(IRM, na.rm = TRUE),
    .groups = "drop"
  )

# -------------------------
# Function to Create Heatmaps
# -------------------------
generate_heatmap <- function(df, metric, title, low_color, mid_color, high_color, limits) {
  ggplot(df, aes(x = factor(Method), y = Protein, fill = .data[[metric]])) +
    geom_tile(color = "white", na.rm = FALSE) +  # Ensure missing data appears white
    scale_fill_gradient2(
      low = low_color, mid = mid_color, high = high_color, 
      midpoint = median(limits), limits = limits, na.value = "white"
    ) +   
    labs(title = "", x = "Method", y = "Protein", fill = title) +  # Removed "Max_"
    theme_minimal() +
    custom_theme
}

# -------------------------
# Generate Individual Heatmaps
# -------------------------
plot_spearman <- generate_heatmap(heatmap_data, "Spearman", "Spearman Correlation", "#F8F9FF", "#839FCC", "#0E3F5C", c(0,1))
plot_pearson  <- generate_heatmap(heatmap_data, "Pearson", "Pearson Correlation", "#F9F9F9", "#B3ABE8", "#312271", c(0,1))
plot_mae      <- generate_heatmap(heatmap_data, "MAE", "MAE", "#F9F7EA", "#B3CD9D", "#006E37", range(heatmap_data$MAE, na.rm = TRUE))
plot_rmse     <- generate_heatmap(heatmap_data, "RMSE", "RMSE", "#FADDC3", "#F69E71", "#EA4C3B", range(heatmap_data$RMSE, na.rm = TRUE))
plot_irm      <- generate_heatmap(heatmap_data, "IRM", "IRM", "#F8DCD9", "#C46EAD", "#583C88", range(heatmap_data$IRM, na.rm = TRUE))

# -------------------------
# Arrange Plots Using gridExtra
# -------------------------
final_plot <- grid.arrange(plot_spearman, plot_pearson, plot_mae, plot_rmse, plot_irm, ncol = 2, nrow = 3)

# -------------------------
# Save the Final Figure
# -------------------------
ggsave("04_Figures/R/Fig6.tiff", plot = final_plot, width = 18, height = 16, dpi = 300)
```

# **Impact of protocol variants on VS performance**

## Fig. S2

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(stringr)
library(gridExtra)  # Alternative to patchwork for combining plots

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(color = "black", size = 18),
  legend.title = element_text(size = 18, face = "bold"),  # Increased legend title size
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 16, angle = 0, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 16),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")  # Increases Protein label size
)

# -------------------------
# Filter Data (IC50, Consider Variants 1-5, DataFusion = Minimum)
# -------------------------
df_ic50 <- df %>%
  filter(
    ExpVariable == "IC50",
    EnergyTerm %in% c("MMGBSABind_Coulomb", "docking.score"),  
    Pose <= 19,  
    DataFusion == "Minimum"
  )

# -------------------------
# Summarize Maximum Metrics per Protein and Variant
# -------------------------
heatmap_data <- df_ic50 %>%
  group_by(Protein, Variant) %>%
  summarise(
    Spearman = max(Spearman, na.rm = TRUE),
    Pearson = max(Pearson, na.rm = TRUE),
    MAE = max(MAE, na.rm = TRUE),
    RMSE = max(RMSE, na.rm = TRUE),
    IRM = max(IRM, na.rm = TRUE),
    .groups = "drop"
  )

# -------------------------
# Function to Create Heatmaps
# -------------------------
generate_heatmap <- function(df, metric, title, low_color, mid_color, high_color, limits) {
  ggplot(df, aes(x = factor(Variant), y = Protein, fill = .data[[metric]])) +
    geom_tile(color = "white", na.rm = FALSE) +  # Ensure missing data appears white
    scale_fill_gradient2(
      low = low_color, mid = mid_color, high = high_color, 
      midpoint = median(limits), limits = limits, na.value = "white"
    ) +   
    labs(title = "", x = "Variant", y = "Protein", fill = title) +  # Removed "Max_"
    theme_minimal() +
    custom_theme
}

# -------------------------
# Generate Individual Heatmaps
# -------------------------
plot_spearman <- generate_heatmap(heatmap_data, "Spearman", "Spearman Correlation", "#F8F9FF", "#839FCC", "#0E3F5C", c(0,1))
plot_pearson  <- generate_heatmap(heatmap_data, "Pearson", "Pearson Correlation", "#F9F9F9", "#B3ABE8", "#312271", c(0,1))
plot_mae      <- generate_heatmap(heatmap_data, "MAE", "MAE", "#F9F7EA", "#B3CD9D", "#006E37", range(heatmap_data$MAE, na.rm = TRUE))
plot_rmse     <- generate_heatmap(heatmap_data, "RMSE", "RMSE", "#FADDC3", "#F69E71", "#EA4C3B", range(heatmap_data$RMSE, na.rm = TRUE))
plot_irm      <- generate_heatmap(heatmap_data, "IRM", "IRM", "#F8DCD9", "#C46EAD", "#583C88", range(heatmap_data$IRM, na.rm = TRUE))

# -------------------------
# Arrange Plots Using gridExtra
# -------------------------
final_plot <- grid.arrange(plot_spearman, plot_pearson, plot_mae, plot_rmse, plot_irm, ncol = 2, nrow = 3)

# -------------------------
# Save the Final Figure
# -------------------------
ggsave("04_Figures/R/FigS2.tiff", plot = final_plot, width = 18, height = 16, dpi = 300)
```

# **Frame dependence across variants 3, 4, and 5**

## Fig. S3

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(stringr)

# Define custom theme for consistency
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(color = "black", size = 18),
  legend.title = element_text(size = 18, face = "bold"),  # Increased legend title size
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 16, angle = 0, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 7),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 18, face = "bold")  # Increases Protein label size
)

# -------------------------
# Filter Data (IC50, Exclude Variants 1 & 2)
# -------------------------
df_ic50 <- df %>%
  filter(ExpVariable == "IC50", 
         !Variant %in% c(1, 2), 
         EnergyTerm %in% c("MMGBSABind_Coulomb", "docking.score"))

# -------------------------
# Process Frame Column (Ensure Global Ordering)
# -------------------------
df_ic50 <- df_ic50 %>%
  mutate(
    # Extract numerical part from Frame (if exists), keep NA otherwise
    Frame_Num = as.numeric(str_extract(Frame, "\\d+")),
    
    # Assign ordering: 0 first, "All_Frames" second, then numerical order
    Frame_Order = case_when(
      Frame == "0" ~ -2,           # Assign lowest order to "0"
      Frame == "All_Frames" ~ -1,  # Assign second lowest order to "All_Frames"
      TRUE ~ Frame_Num             # Otherwise, use extracted numerical value
    )
  )

# -------------------------
# Define Global Frame Order
# -------------------------
global_frame_levels <- df_ic50 %>%
  arrange(Frame_Order) %>%
  distinct(Frame) %>%
  pull(Frame)  # Extract ordered unique Frame values

# Apply the global frame order across all proteins
df_ic50 <- df_ic50 %>%
  arrange(Protein, Frame_Order) %>%
  mutate(Frame = factor(Frame, levels = global_frame_levels)) %>%
  select(-Frame_Num, -Frame_Order)  # Remove helper columns

# -------------------------
# Summarize Spearman Correlation per Frame, Variant, and Protein
# -------------------------
heatmap_data <- df_ic50 %>%
  group_by(Protein, Frame, Variant) %>%
  summarise(Mean_Spearman = max(Spearman, na.rm = TRUE), .groups = "drop")

# -------------------------
# Ensure Specific Frames Exist for Variant 4 in 6QSU
# -------------------------
excluded_frames <- c("frame1000", "frame978", "frame901", "frame861", "frame801")

missing_rows <- expand.grid(
  Protein = "6QSU",
  Frame = excluded_frames,
  Variant = 4,
  Mean_Spearman = NA
)

heatmap_data <- bind_rows(heatmap_data, missing_rows)  # Add missing frames

# -------------------------
# Generate Heatmap Faceted by Protein (Fix Frame Ordering)
# -------------------------
heatmap_plot <- ggplot(heatmap_data, aes(x = Variant, y = Frame, fill = Mean_Spearman)) +
  geom_tile(color = "grey80", na.rm = FALSE) +  # Keep grid visible with grey lines
  scale_fill_gradient2(low = "#4F53B7", mid = "yellow", high = "#005600", midpoint = 0, 
                       na.value = "white", limits = c(-1, 1)) +  # Force legend limits from -1 to 1
  labs(title = "",
       x = "Variant",
       y = "Frames",
       fill = "Spearman correlation") +
  theme_minimal() +
  custom_theme +
  facet_wrap(~ Protein, scales = "free_y")  # Allow different Frame scales per Protein

# Display the plot
heatmap_plot

# -------------------------
# Save Heatmap as TIFF
# -------------------------
ggsave("04_Figures/R/FigS3.tiff", plot = heatmap_plot, width = 14, height = 12, dpi = 300)
```

# **Comparison of IC₅₀ and pIC₅₀ as experimental parameters for correlation analysis**

## Fig. S4

```{r}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(gridExtra)  # For combining multiple plots

# -------------------------
# Define Custom Theme for Consistency
# -------------------------
custom_theme <- theme(
  legend.position = "top",
  legend.key.size = unit(1.5, "cm"),
  legend.text = element_text(color = "black", size = 18),
  legend.title = element_text(size = 20, face = "bold"),
  axis.title = element_text(color = "black", face = "bold", size = 18),
  axis.text.x = element_text(colour = "black", face = "bold", size = 16, angle = 0, hjust = 1),
  axis.text.y = element_text(colour = "black", face = "bold", size = 16),
  axis.ticks = element_line(colour = "black", size = 0.5),
  panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
  strip.text = element_text(size = 22, face = "bold", vjust = 1)  # Bigger titles, closer to plots
)

# -------------------------
# Filter Data (IC50 and pIC50)
# -------------------------
df_comparison <- df %>%
  filter(ExpVariable %in% c("IC50", "pIC50"),
         EnergyTerm %in% c("MMGBSABind_Coulomb"),
         Pose <= 19)

# -------------------------
# Summarize Spearman and Pearson Correlation per Protein and ExpVariable
# -------------------------
correlation_data <- df_comparison %>%
  group_by(Protein, DataFusion, ExpVariable) %>%
  summarise(
    Spearman = median(Spearman, na.rm = TRUE),
    Pearson = median(Pearson, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = ExpVariable, values_from = c(Spearman, Pearson)) %>%
  mutate(
    Spearman = (abs(Spearman_IC50) - abs(Spearman_pIC50)),  
    Pearson  = (abs(Pearson_IC50) - abs(Pearson_pIC50))     
  )

# -------------------------
# Reshape for Visualization
# -------------------------
heatmap_data <- correlation_data %>%
  select(Protein, DataFusion, Spearman, Pearson) %>%
  pivot_longer(cols = c(Spearman, Pearson), names_to = "Correlation Type", values_to = "Diff")

# -------------------------
# Function to Generate Heatmaps Per Protein
# -------------------------
generate_heatmap <- function(protein_name, df) {
  ggplot(df, aes(x = `Correlation Type`, y = DataFusion, fill = Diff)) +
    geom_tile(color = "white", na.rm = FALSE) +  
    scale_fill_gradientn(
      colors = c("#9E3D22", "#EF8737", "#B0DEA5", "#699CC5", "#2B5C8A"),
      values = scales::rescale(c(-0.3, -0.15, 0, 0.15, 0.3)),
      limits = c(-0.3, 0.3),
      na.value = "white",
      guide = guide_colorbar(
        barwidth = 15, barheight = 1.8, title.position = "top", title.hjust = 0.5,
        label.theme = element_text(size = 16, face = "bold"),
        title.theme = element_text(size = 20, face = "bold"),
        frame.colour = "black"
      )
    ) +
    labs(
      title = protein_name,  # Direct title
      x = "",
      y = "Data Fusion Method",
      fill = "Difference"
    ) +
    theme_minimal() +
    custom_theme +
    theme(
      plot.title = element_text(size = 24, face = "bold", hjust = 0.5, vjust = 0.2),  # Bigger title, closer
      legend.text = element_text(size = 16, face = "bold"),
      legend.title = element_text(size = 18, face = "bold")
    )
}

# -------------------------
# Generate Heatmaps for Each Protein
# -------------------------
protein_list <- unique(heatmap_data$Protein)

plots <- lapply(protein_list, function(protein) {
  generate_heatmap(protein, heatmap_data %>% filter(Protein == protein))
})

# -------------------------
# Arrange All Heatmaps in a Single Figure
# -------------------------
final_plot <- grid.arrange(grobs = plots, ncol = 2, nrow = 2)

# -------------------------
# Save the Final Figure
# -------------------------
ggsave("04_Figures/R/FigS4.tiff", plot = final_plot, width = 16, height = 12, dpi = 300)

# Display final plot
print(final_plot)

```

# Frequency Analysis of high-correlation parameter combinations in VS

## Table S2

```{r}

library(dplyr)

# -------------------------
# Step 2: Filter Data for High Spearman Correlations
# -------------------------
df_high_spearman <- df %>%
  filter(Spearman >= 0.75, ExpVariable == "IC50") %>%
  select( DataFusion, EnergyTerm, Method, Variant, Spearman, Pearson, MAE, RMSE)  # Keep Spearman for averaging

# -------------------------
# Step 3: Count Frequent Combinations and Compute Average Spearman
# -------------------------
frequent_combinations <- df_high_spearman %>%
  group_by(DataFusion, EnergyTerm, Method, Variant) %>%
  summarise(
    Frequency = n(),  # Count occurrences
    "Avg. Spearman" = mean(Spearman, na.rm = TRUE),  # Compute average Spearman
    "Avg. Pearson" = mean(Pearson, na.rm = TRUE),
    "Avg. MAE" = mean(MAE, na.rm = TRUE),
    "Avg. RMSE" = mean(RMSE, na.rm = TRUE),
    #Avg_IRM = mean(IRM, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Frequency))




frequent_combinations2 <- frequent_combinations %>%
  filter(Frequency >= 50)

# -------------------------
# Step 4: Display Data
# -------------------------
print(frequent_combinations2)
```

## Sankey (optional)

```{r}

library(networkD3)
library(dplyr)
library(tidyr)

# -------------------------
# Step 1: Filter Frequent Combinations for Sankey
# -------------------------
df_sankey <- frequent_combinations2 %>%
  mutate(Variant = as.character(Variant)) %>%  # Convert Variant to character
  select(DataFusion, EnergyTerm, Method, Variant, Frequency) %>%
  pivot_longer(cols = c(DataFusion, EnergyTerm, Method, Variant), names_to = "source", values_to = "target")

# -------------------------
# Step 2: Count Occurrences
# -------------------------
df_sankey_count <- df_sankey %>%
  group_by(source, target) %>%
  summarise(n = sum(Frequency), .groups = "drop") %>%
  arrange(desc(n))

# -------------------------
# Step 3: Convert Nodes to Index Format
# -------------------------
nodes <- data.frame(name = unique(c(df_sankey_count$source, df_sankey_count$target)))

# -------------------------
# Step 4: Assign Group Colors to Nodes
# -------------------------
nodes$group <- case_when(
  nodes$name %in% df_sankey$source[df_sankey$source == "Method"] ~ "Method",
  nodes$name %in% df_sankey$source[df_sankey$source == "EnergyTerm"] ~ "EnergyTerm",
  nodes$name %in% df_sankey$source[df_sankey$source == "DataFusion"] ~ "DataFusion",
  nodes$name %in% df_sankey$source[df_sankey$source == "Variant"] ~ "Variant",
  TRUE ~ "Inherited"  # Default category for targets
)

# -------------------------
# Step 5: Define Custom Colors for Source-Target Mapping
# -------------------------
color_scale <- 'd3.scaleOrdinal()
  .domain(["Method", "EnergyTerm", "DataFusion", "Variant", "Inherited"])
  .range(["#D73027", "#4575B4", "#1A9850", "#F46D43", "#999999"])'  # Targets inherit faded colors

# -------------------------
# Step 6: Convert Source and Target to Numeric Indexes
# -------------------------
df_sankey_count$source <- match(df_sankey_count$source, nodes$name) - 1
df_sankey_count$target <- match(df_sankey_count$target, nodes$name) - 1

# -------------------------
# Step 7: Generate Sankey Diagram with Color Control
# -------------------------
sankey <- sankeyNetwork(
  Links = df_sankey_count, Nodes = nodes,
  Source = "source", Target = "target",
  Value = "n", NodeID = "name",
  fontSize = 14, nodeWidth = 30,
  NodeGroup = "group",  # Assign colors based on group
  colourScale = JS(color_scale),  # Apply color scale
  sinksRight = FALSE  # Ensure target nodes get color
)

# -------------------------
# Step 8: Display Sankey Diagram
# -------------------------
sankey
```
